### 2021-07-08 supplementary/realdata-cust2-orig.Rmd

---
title: "Clustering real data with two customers"
output:
  bookdown::pdf_book:
    MonashEBSTemplates::workingpaper:
    #base_format: rticles::asa_article
    fig_height: 8
    fig_width: 12
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
read_chunk("R/simdata01.R")
```

```{r load-lib}

```

# What you are doing?

Taking real customers from the smart meter data set and cluster them and see if their distributions are different

-  just one cyclic granularity (hod)
-  two cyclic granularities (hod and dow)

_Compute quantiles of conditional distributions_
Conditional quantiles are obtained for each combination of categories.

## entire data
```{r}
sm <- gravitas::smart_meter10 %>%
  filter(customer_id %in% c("10006704", "10017936"))

sm
```

## how do they look

```{r}
sm %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id, ncol = 1) +
  theme_bw()
```

```{r}
sm %>% 
  create_gran("hour_day") %>% 
  ggplot(aes(x = hour_day, y = general_supply_kwh)) + 
  geom_boxplot() + 
  facet_wrap(~customer_id, ncol = 1) 
```

## make one granularity

```{r}
sm_hod <- gravitas::create_gran(sm, "hour_day")
```

## conditional list across categories of one granularity

```{r}
sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

```

## conditional quantiles across categories of one granularity

```{r}
quantile_prob_val = seq(0.1, 0.9, 0.1)

ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

```

## how do their deciles look for different categories of cyclic granularity

```{r}
categories_ref <- names(sm_hod_list)[-1] %>% as_tibble() %>% 
  mutate(categories_serial_id = as.character(row_number()))

data_heatmap <- sm_hod_quantiles %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  left_join(categories_ref) %>% 
  rename("category" = "value") %>% 
  select(-categories_serial_id) %>% 
  mutate(category = as.numeric(category))

data_heatmap%>% 
ggplot(aes(x= customer_serial_id, y = as.factor(quantile_prob))) +
  geom_tile(aes(fill = quantile_val)) +
  scale_fill_distiller(palette = "YlGnBu")+ 
  facet_wrap(~category) +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Deciles") +
  xlab("data sets")
```

```{r, eval = FALSE}
data_heatmap %>% 
  ggplot() + geom_line(aes(x=category, y = quantile_val, color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1) + 
  scale_color_brewer(palette = "Dark2") + 
  #scale_y_log10() +
  theme_bw()

data_heatmap %>% 
  pivot_wider(names_from = customer_serial_id, values_from = quantile_val) %>% 
  group_by(quantile_prob) %>% 
  mutate(percent_diff = if_else(`1`==0, 0, (`2`- `1`)*100/`1`)) %>%  ggplot() + geom_line(aes(x=category, y = percent_diff, color = quantile_prob)) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
```

## Iterations of each customer

### Data preparation

```{r}
sample_seed <-  seq(100, 500, 100)
ncust <- sm %>% distinct(customer_id)

all_data_clust = map_dfr(seq_len(nrow(ncust)), function(x){
  map(seq_len(length(sample_seed)), function(seed){
     set.seed(sample_seed[seed])
     data <- sm %>% as_tibble %>% filter(customer_id %in% ncust$customer_id[x])
     nr = nrow(data)

     data1 <- data %>% 
     mutate(customer_id = paste(customer_id, seed, sep = "-"),
             general_supply_kwh = general_supply_kwh + rnorm(1, 0, 0.01))
     data1
}) %>% bind_rows(.id = "seed_id")
}) %>% bind_rows(sm %>% as_tibble() %>% bind_cols(seed_id = "0")) %>% 
  tsibble::as_tsibble(index = reading_datetime, key  = customer_id)
```

```{r}
all_data_clust %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id, ncol = 1) +
  theme_bw()
```

### make one granularity

```{r}
sm_hod <- gravitas::create_gran(all_data_clust, "hour_day")
```

## conditional list across categories of one granularity

```{r}
sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

lookup_tbl <- sm_hod_list %>% 
  mutate(
         customer_serial_id = row_number()) %>% 
  select(customer_id, customer_serial_id)

```

## conditional quantiles across categories of one granularity

```{r}
quantile_prob_val = seq(0.1, 0.9, 0.1)

ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

```

## JS Pairwise distances between households

```{r}
sm_dist_data <- sm_hod_quantiles %>% 
  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, values_from = data)
```

```{r JS}

```

```{r}
#ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
nrow_data <- nrow(sm_dist_data)
ncol_data <-  ncol(sm_dist_data[-1])

dist_data <- map(seq_len(nrow_data), function(x){ # first data
  map(seq_len(nrow_data), function(y){ # 2nd data
    map(seq_len(ncol_data), function(z){ # number of combinations nx*nfacet
      JS(
        prob = quantile_prob_val,
        unlist(sm_dist_data[-1] %>% extract(x, z)),
        unlist(sm_dist_data[-1] %>% extract(y, z))
      ) %>% as_tibble()
    })%>% bind_rows(.id = "combinations")
  })%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

dist_mat <- dist_data %>%
  mutate(data_type1 = as.numeric(data_type1), data_type2 = as.numeric(data_type2)) %>% 
  group_by(data_type1, data_type2) %>%
  summarise(dist = sum(value)) %>%
  pivot_wider(names_from = data_type2,
              values_from = dist)

dist_mat_format <- dist_mat %>%
  ungroup() %>% 
  select(-data_type1)

rownames(dist_mat_format) <- lookup_tbl$customer_id

```

### hc

```{r}
d = stats::as.dist(dist_mat_format)
hc = stats::hclust(d,method="complete")
plot(hc)

mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- lookup_tbl$customer_id

groups<-cutree(hc, k=2)

all_data_cluster <- cbind(dist_mat_format, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups)) %>% as_tibble()

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups,
       label = rownames(mds))) +
  geom_point(size = 1) +
  geom_text(check_overlap = TRUE)  +
  theme_classic()+
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Dark2")

mds_plot_10
```

# Scaled data

### make one granularity

```{r}
sm_hod <- gravitas::create_gran(all_data_clust, "hour_day")
```

## conditional list across categories of one granularity

```{r}

sm_hod <- sm_hod %>% 
  group_by(customer_id) %>% 
  mutate(general_supply_kwh = scale(general_supply_kwh))

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

lookup_tbl <- sm_hod_list %>% 
  mutate(
         customer_serial_id = row_number()) %>% 
  select(customer_id, customer_serial_id)

```

## conditional quantiles across categories of one granularity

```{r}
quantile_prob_val = seq(0.1, 0.9, 0.1)

ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

```

## JS Pairwise distances between households

```{r}
sm_dist_data <- sm_hod_quantiles %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, values_from = data)
```

```{r}
#ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
nrow_data <- nrow(sm_dist_data)
ncol_data <-  ncol(sm_dist_data[-1])

dist_data <- map(seq_len(nrow_data), function(x){ # first data
  map(seq_len(nrow_data), function(y){ # 2nd data
    map(seq_len(ncol_data), function(z){ # number of combinations nx*nfacet
      JS(
        prob = quantile_prob_val,
        unlist(sm_dist_data[-1] %>% extract(x, z)),
        unlist(sm_dist_data[-1] %>% extract(y, z))
      ) %>% as_tibble()
    })%>% bind_rows(.id = "combinations")
  })%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

dist_mat <- dist_data %>%
  mutate(data_type1 = as.numeric(data_type1), data_type2 = as.numeric(data_type2))%>%
  group_by(data_type1, data_type2) %>%
  summarise(dist = sum(value)) %>%
  pivot_wider(names_from = data_type2,
              values_from = dist)

dist_mat_format <- dist_mat %>%
  ungroup() %>% 
  select(-data_type1)

rownames(dist_mat_format) <- lookup_tbl$customer_id

```

### hc

```{r}
d = stats::as.dist(dist_mat_format)
hc = stats::hclust(d,method="complete")
plot(hc)
```

```{r}
mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- lookup_tbl$customer_id

library(fpc)
koptimal = fpc::nselectboot(as.dist(dist_mat_format),
                        B = 50,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:12)

koptimal = koptimal$kopt

groups<-cutree(hc, k=koptimal)

all_data_cluster <- cbind(dist_mat_format, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups)) %>% as_tibble()

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups,
       label = rownames(mds))) +
  geom_point(size = 1) +
  geom_text(check_overlap = TRUE)  +
  theme_classic()+
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Dark2")

mds_plot_10
```

_JS Pairwise distances between datasets_
Distance between the data sets is computed as the sum of JS distances across different categories.

_Hierarchical clustering with 4 clusters_
Hierarchical clustering is performed using k = 4 and dendogram observed

_Clusters obtained visualized using MDS_
Each cluster represents data sets from a separate design