### 2021-07-16 old/simdata_cluster02.Rmd

---
title: "Clustering"
author: "Sayani Gupta"
date: "16/06/2021"
output:
  bookdown::pdf_book:
    MonashEBSTemplates::workingpaper:
    #base_format: rticles::asa_article
    fig_height: 8
    fig_width: 12
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
read_chunk("simdata01.R")
```

# How the time series plot of the designs look like for different sample sizes?

```{r sim-val, echo = TRUE}
set.seed(9999)
nx_val = 2 # number of x-axis levels
nfacet_val = 3 # number of facet levels
w1_val = 3 # increment in mean
w2_val = 0 # increment in sd
mean_val = 0 # mean of normal distribution of starting combination
sd_val = 1 # sd of normal distribution of starting combination
quantile_prob_val = seq(0.1, 0.9, 0.1)
```

```{r load-lib}
```

```{r designs}
```

## Number of obervations per combination is 4

```{r, echo = TRUE}
ntimes_val = 4
```

```{r data}
```

```{r category-plots}
```

```{r}
change_index <- function(data){

index_new <- map(seq_len(ntimes_val), function(i){
  map((seq_len(nx_val*nfacet_val)), function(j)
    {
value = i + (j-1)*ntimes_val
})
  }) %>% unlist()

data_new = data %>%
  ungroup() %>%
  mutate(index_old = row_number(),
         index_new = index_new)

y = data_new[match(index_new, data_new$index_old),]

y <- y %>%
  mutate(time = row_number())

return(y)
}

endbreaks<- nrow(sim_panel_null)

p1 <- change_index(sim_panel_null) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p2 <- change_index(sim_panel_varf) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p3 <- change_index(sim_panel_varx) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue")+
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p4 <- change_index(sim_panel_varall) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
  panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

t1 <- sim_panel_varall %>% ungroup() %>%  select (-c(1, 2))

# %>% 
#   kable(format = "latex", caption = "Simulated data for categories format")

t2 <- change_index(sim_panel_varall) %>% select (-c(1, 2))

# %>% 
#   kable(format = "latex",
#         caption = "Simulated data for time series format")

kable(list(t1, t2), format = "latex", caption = "Simulated data across combination of categories (left) and manipulated time series format (right)") %>% 
  row_spec(1:24,  background =  "#d3d3d3")

# kable(t1) %>%
#   kable_styling(full_width = FALSE, position = "float_left")
# kable(t2) %>%
#   kable_styling(full_width = FALSE, position = "left")
```

```{r}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
```

## Number of obervations per combination is 10

```{r nobs10, echo = TRUE}
ntimes_val = 10 # number of observations per combination 
```

```{r data}
```

```{r category-plots}
```

```{r line-graph}
```

```{r plot10}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)

```

## Number of obervations per combination is 50

```{r nobs50, echo = TRUE}
ntimes_val = 50 # number of observations per combination 
```

```{r data}
```

```{r category-plots}
```

```{r line-graph}
```

```{r plot50}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)

```

# Simulation: 4 datasets with 2 designs

Generate 4 series, two from design varf and another two from varall and see if clustering happens properly through the approaches

## Generate 2 time series from varx

```{r}
set.seed(9999)
```

```{r data_varf, echo = TRUE}

data1_varf <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varf_normal(2, 3, 0, 1, 5, 0)
) %>% unnest(data)

data2_varf <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varf_normal(2, 3, 0, 1, 2, 0)
) %>% unnest(data)
```

## Generate 2 time series from varall

```{r data_varall, echo = TRUE}
data1_varall <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varall_normal(2, 3, 0, 1, 5, 0)
) %>% unnest(data)

data2_varall <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varall_normal(2, 3, 0, 1, 2, 0)
) %>% unnest(data)
```

## How do they look?

```{r}
p_data1_varf <- data1_varf %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) +
  facet_wrap(~id_facet) +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(b)", round(varf, 2))) +
  xlab("x level")+
  theme_bw() +
  ggtitle("varf-1")

p_data2_varf <- data2_varf %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) +
  facet_wrap(~id_facet) +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(b)", round(varf, 2))) +
  xlab("x level")+
  theme_bw()+
  ggtitle("varf-2")

p_data1_varall <- data1_varall %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) +
  facet_wrap(~id_facet) +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(b)", round(varf, 2))) +
  xlab("x level")+
  theme_bw()+
  ggtitle("varall-1")

p_data2_varall <- data2_varall %>%
  ggplot(aes(x = as.factor(id_x), y = sim_data)) +
  facet_wrap(~id_facet) +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(b)", round(varf, 2))) +
  xlab("x level")+
  theme_bw()+
  ggtitle("varall-2")

(p_data1_varf +p_data2_varf)/(p_data1_varall+p_data2_varall)
```

## Compute quantiles of conditional distributions

```{r comp_quant}

compute_quantiles(data1_varf) %>%
   unnest(sim_data_quantile) %>%
   ggplot() +
   geom_boxplot(aes(x = as.factor(id_x), y = sim_data_quantile)) +
   facet_wrap(~id_facet)

data_q <- bind_rows(data1_varf = 
compute_quantiles(data1_varf, 
                                 quantile_prob = quantile_prob_val) ,

 data2_varf = compute_quantiles(data2_varf,
                                 quantile_prob = quantile_prob_val),

 data1_varall = compute_quantiles(data1_varall,
                                 quantile_prob = quantile_prob_val) ,

 data2_varall = compute_quantiles(data2_varall,
                                 quantile_prob = quantile_prob_val), .id = "data_type"
)

#%>% mutate(data_type = as.integer(data_type))

```

## JS Pairwise distances between data sets

```{r JS}
```

```{r JS-pairwise, echo=TRUE}
data_q_wide <- data_q %>% 
  pivot_wider(names_from = data_type,
              values_from = sim_data_quantile) %>% 
  select(-c(1,2))

ndata <- data_q %>% distinct(data_type) %>% pull(data_type)
ldata <- length(ndata)
lcomb <-  nx_val*nfacet_val

dist_data <- map(1:ldata, function(x){ # first data
  map(1:ldata, function(y){ # 2nd data
    map(1:lcomb, function(z){ # number of combinations nx*nfacet
    JS(
        prob = quantile_prob_val,
        unlist(data_q_wide[z,x]),
        unlist(data_q_wide[z,y])
      ) %>% as_tibble()
  })%>% bind_rows(.id = "combinations")
})%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

```

## Hierarchical clustering based on pairwise distances

```{r hc}

dist_mat <- dist_data %>% 
  group_by(data_type1, data_type2) %>% 
  summarise(dist = sum(value)) %>% 
  pivot_wider(names_from = data_type2,
              values_from = dist) %>% 
  ungroup() %>% 
  select(-1)

d = stats::as.dist(dist_mat)

hc = stats::hclust(d,method="complete")
plot(hc)
```

## Multi-dimensional scaling with hierarchical clusters

```{r mds-plot}

mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")

mds <- mds %>%
  mutate(data_type = ndata) %>%
  select((data_type)
         , Dim.1, Dim.2)

groups<-cutree(hc, k=2)

all_data_cluster <- cbind(dist_mat, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups))

mds_plot <- ggpubr::ggscatter(all_data_cluster,
                        x = "Dim.1",
                        y = "Dim.2",
                        label = mds$data_type,
                        color = "groups",
                        repel = TRUE,
                        size = 2) +
  scale_color_brewer(palette = "Dark2")
```

```{r}
p1 <- change_index(data1_varf) %>%
  ggplot(aes(x = index_new,
             y = sim_data)) +
  geom_line() +
  #scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue")
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("")+
  xlab("index") +
  ggtitle("data1_varf")

p2 <- change_index(data2_varf) %>%
  ggplot(aes(x = index_new,
             y = sim_data)) +
  geom_line() +
  #scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("")+
  xlab("index")+
  ggtitle("data2_varf")

p3 <- change_index(data1_varall) %>%
  ggplot(aes(x = index_new,
             y = sim_data)) +
  geom_line() +
  #scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue")+
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("")+
  xlab("index")+
  ggtitle("data1_varall")

p4 <- change_index(data2_varall) %>%
  ggplot(aes(x = index_new,
             y = sim_data)) +
  geom_line() +
  #scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
  panel.grid.minor.x =  element_blank())+
  ylab("")+
  xlab("index")+
  ggtitle("data2_varall")

mds_plot+(p1/p2/p3/p4)
```

# Repeat with 3 designs (for varying sample size)

varf and varall - 100 obs each combination and varx with 10 and 50 observation per combination taken

```{r makedata}
data1_varx <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = 10,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varx_normal(2, 3, 0, 1, 1.5, 0)
) %>% unnest(data)

data2_varx <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = 50,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varx_normal(2, 3, 0, 1, 10, 0)
) %>% unnest(data)

# make-quantiles
data_q <- bind_rows(data1_varf = 
compute_quantiles(data1_varf, 
                                 quantile_prob = quantile_prob_val) ,

 data2_varf = compute_quantiles(data2_varf,
                                 quantile_prob = quantile_prob_val),

 data1_varall = compute_quantiles(data1_varall,
                                 quantile_prob = quantile_prob_val) ,

 data2_varall = compute_quantiles(data2_varall,
                                 quantile_prob = quantile_prob_val),

 data1_varx = compute_quantiles(data1_varx,
                                 quantile_prob = quantile_prob_val) ,

 data2_varx = compute_quantiles(data2_varx,
                                 quantile_prob = quantile_prob_val),

.id = "data_type"
)

# Js-pairwise

data_q_wide <- data_q %>% 
  pivot_wider(names_from = data_type,
              values_from = sim_data_quantile) %>% 
  select(-c(1,2))

ndata <- data_q %>% distinct(data_type) %>% pull(data_type)
ldata <- length(ndata)
lcomb <-  nx_val*nfacet_val

dist_data <- map(1:ldata, function(x){ # first data
  map(1:ldata, function(y){ # 2nd data
    map(1:lcomb, function(z){ # number of combinations nx*nfacet
    JS(
        prob = quantile_prob_val,
        unlist(data_q_wide[z,x]),
        unlist(data_q_wide[z,y])
      ) %>% as_tibble()
  })%>% bind_rows(.id = "combinations")
})%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

# hc

dist_mat <- dist_data %>% 
  group_by(data_type1, data_type2) %>% 
  summarise(dist = sum(value)) %>% 
  pivot_wider(names_from = data_type2,
              values_from = dist) %>% 
  ungroup() %>% 
  select(-1)

rownames(dist_mat) <- ndata
d = stats::as.dist(dist_mat)

hc = stats::hclust(d,method="complete")
plot(hc)

# mds

mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")

mds <- mds %>%
  mutate(data_type = ndata) %>%
  select((data_type)
         , Dim.1, Dim.2)

groups<-cutree(hc, k=3)

all_data_cluster <- cbind(dist_mat, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups))

mds_plot <- ggpubr::ggscatter(all_data_cluster,
                        x = "Dim.1",
                        y = "Dim.2",
                        label = mds$data_type,
                        color = "groups",
                        repel = TRUE,
                        size = 2,
                        shape = 1) 
mds_plot

```

<!-- # Repeat 4 designs each 10 times and see if there are 4 clusters -->

<!-- ```{r} -->

<!-- sample_seed <- seq(10, 100, 10) -->

<!-- data_varall <- map(1:10, function(x){ -->
<!-- set.seed(sample_seed[x])   -->
<!-- sim_panel_varall <- sim_panel( -->
<!--   nx = nx_val, nfacet =  nfacet_val, -->
<!--   ntimes = 100, -->
<!--   # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5) -->
<!--   sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val) -->
<!-- ) %>% unnest(data) -->
<!-- }) %>% bind_rows(.id = "seed_id") %>%  -->
<!--   mutate(date_type = "varall") -->

<!-- change_index(data_varall, ntimes_val = 5, nx_val = 2, nfacet_val = 3) -->

<!-- ``` -->