### 2021-07-16 supplementary/realdata-cust100-noiter.Rmd

---
title: "Clustering real data with 100 customers (no missing value)"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false
    cache: TRUE
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
read_chunk("R/simdata01.R")
```

```{r load-lib}
```

# Clutering approach

1. _Compute quantiles of distributions across each hour of day_
2. _Compute JS distance between households for each hour of day_
3. _Total distance between households computed as sum of JS distances for all hours_
4. _Cluster using this distance with hierarchical clustering algorithm (method "complete")_

```{r read-data}
# elec_ts <- read_rds("../data/elec_ts.rds")
# 
# set.seed(12345)
# 
# sm_100 <- elec_ts %>%
#   as_tibble() %>% 
#   distinct(customer_id) %>%
#   slice_sample(n = 100)
# 
# sm <- elec_ts %>% 
#   filter(customer_id %in% sm_100$customer_id) %>% 
#   ungroup()

#elec_nogap <- read_rds("../data/elec_nogap.rds")
```

```{r read100}
# 
# set.seed(12345)
# 
# sm_100 <- elec_nogap %>%
#   as_tibble() %>%
#   distinct(customer_id) %>%
#   slice_sample(n = 100)
# 
# sm <- elec_nogap %>%
#   filter(customer_id %in% sm_100$customer_id) %>%
#   ungroup()
# write_rds(sm, "../data/elec_nogap-100cust.rds")

sm <- read_rds("../data/elec_nogap-100cust.rds")
```

```{r hod}
sm_hod <- gravitas::create_gran(sm, "hour_day") 
```

```{r}
sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

```

```{r heatmpa-data}
quantile_prob_val = seq(0.1, 0.9, 0.1)

ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id") %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id))

categories_ref <- names(sm_hod_list)[-1] %>% as_tibble() %>% 
  mutate(categories_serial_id = row_number())

data_heatmap <- sm_hod_quantiles %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  left_join(categories_ref) %>% 
  rename("category" = "value") %>% 
  select(-categories_serial_id) %>% 
  mutate(category = as.numeric(category))

```

```{r hod-list}
sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

lookup_tbl <- sm_hod_list %>% 
  mutate(
         customer_serial_id = row_number()) %>% 
  select(customer_id, customer_serial_id)

```

```{r hod-quantile}
quantile_prob_val = seq(0.1, 0.9, 0.1)

ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)]
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list[-1] %>% extract(x, y) %>% unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

sm_filter_nozerosum <- sm_hod_quantiles %>% 
  pivot_longer(-c(1, 2), names_to = "quantiles", values_to = "quantvalue") %>% 
  group_by(customer_serial_id) %>% 
  summarise(m = sum(quantvalue)) %>% 
  filter(m!=0)

sm_hod_quantiles <- sm_hod_quantiles %>%
  filter(customer_serial_id %in% sm_filter_nozerosum$customer_serial_id)

```

```{r hod-dist}
sm_dist_data <- sm_hod_quantiles %>% 
  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, 
           categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, 
              values_from = data) %>% 
  ungroup()
```

```{r JS}

```

```{r hod-dist-mat}
#ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
nrow_data <- nrow(sm_dist_data)
ncol_data <-  ncol(sm_dist_data[-1])

dist_data <- map(seq_len(nrow_data), function(x){ # first data
  map(seq_len(nrow_data), function(y){ # 2nd data
    map(seq_len(ncol_data), function(z){ # number of combinations nx*nfacet
      JS(
        prob = quantile_prob_val,
        unlist(sm_dist_data[-1] %>% extract(x, z)),
        unlist(sm_dist_data[-1] %>% extract(y, z))
      ) %>% as_tibble()
    })%>% bind_rows(.id = "combinations")
  })%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

dist_mat <- dist_data %>%
  mutate(data_type1 = as.numeric(data_type1),
         data_type2 = as.numeric(data_type2)) %>% 
  group_by(data_type1, data_type2) %>%
  summarise(dist = sum(value)) %>%
  pivot_wider(names_from = data_type2,
              values_from = dist)

dist_mat_format <- dist_mat %>%
  ungroup() %>% 
  select(-data_type1)

lookup_tbl <- sm_hod_list %>% 
  mutate(customer_serial_id = row_number()) %>% 
  select(customer_id, customer_serial_id) %>% 
  filter(customer_serial_id %in% sm_filter_nozerosum$customer_serial_id)

rownames(dist_mat_format) <- lookup_tbl$customer_id

```

# Clustering results

```{r hc}
d = stats::as.dist(dist_mat_format)
hc = stats::hclust(d,method="complete")
plot(hc)
```

```{r mds-new}
set.seed(12345)
mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- lookup_tbl$customer_id

library(fpc)
koptimal = fpc::nselectboot(as.dist(dist_mat_format),
                        B = 500,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:15)

koptimal = koptimal$kopt

groups<-cutree(hc, k=koptimal)

all_data_cluster <- cbind(dist_mat_format, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups)) %>% 
  as_tibble()

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups,
       label = rownames(mds))) +
  geom_point(size = 1) +
  geom_text(check_overlap = TRUE)  +
  theme_classic()+
  theme(legend.position = "bottom")

mds_plot_10
```
Optimal number of clusters as defined by the _fpc::nselectboot_ is `r koptimal`.

_k = 15_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 15, border = 2:20)
```

# Cluster characterization

The quartile deviation of different clusters are drawn. The shape of daily load curve is different for all of them. Across each cluster, customers should show different shape and within each cluster, customers should differ in size.

```{r hod-heatmap, fig.cap="Heatmaps show deciles on the y-axis with customers on the x-axis with colors filled by the value of deciles and faceted by hours of the day. It again seems like all the households have gradual change in colors (almost for all hours) as we move to higher deciles except for the 4th household. "}
quantile_prob_val = c(0.1, 0.25, 0.5, 0.75, 0.9)

groups_ref <- groups %>% as_tibble %>% 
              mutate(customer_id = lookup_tbl$customer_id,
                     customer_serial_id = lookup_tbl$customer_serial_id) %>% 
             rename ("group" = "value")

groups_g3 <- groups_ref %>% 
  group_by(group) %>% 
  count() %>% 
  filter(n>=3)

sm_hod_list_cat <- sm_hod_list %>% left_join(groups_ref) %>%
  filter(customer_serial_id %in% sm_filter_nozerosum$customer_serial_id) %>% 
select(group, customer_serial_id, everything())

ncol_sm <- seq_len(ncol(sm_hod_list_cat[-c(1, 2, 3)]))
nrow_sm <- unique(sm_hod_list_cat$group)

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list_cat%>% 
     dplyr::filter(group == x) %>% 
     select(-c(1, 2, 3)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "group_id") %>% 
  mutate(group_id = as.integer(group_id),
         categories_serial_id = as.integer(categories_serial_id))

categories_ref_cat <- names(sm_hod_list_cat)[-c(1, 2, 3)] %>% as_tibble() %>% 
  mutate(categories_serial_id = row_number())

data_heatmap <- sm_hod_quantiles_cat %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  left_join(categories_ref_cat) %>% 
  rename("category" = "value") %>% 
  select(-categories_serial_id) %>% 
  mutate(category = as.numeric(category))

```

```{r cluster-characterization}
data_heatmap %>% 
  filter(group_id %in% groups_g3$group) %>% 
  pivot_wider( names_from = quantile_prob, 
               values_from = quantile_val) %>% 
  ggplot(aes(x = category)) +  
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "#D55E00", alpha = 0.8) +
  geom_ribbon(aes(ymin = `10%`, 
                  ymax = `90%`), fill = "#56B4E9", alpha = 0.4) + 
  geom_line(aes(y = `50%`)) +
  facet_wrap(~group_id, ncol = 2,
             scales = "free_y",
             strip.position = "right") +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
```