### 2021-08-12 sim_setup_algo2_diffmeans.Rmd
adding the handpicking examples

---
title: "Simulation with algorithm 2- different means and one nuisance granularity"
author: "Sayani Gupta"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
library(patchwork)
library(distributional)
library(ggplot2)
library(tidyverse)
library(knitr)
library(readr)
library(GGally)
```

# _Simulation design_

Three circular granularities g1, g2 and g3 are considered with levels 2, 3 and 4 respectively. Many time series with 600 observations are created using the four designs below, each of which is iterated five times. We anticipate to have four clusters, each with five time series conforming to the same design, once we execute the clustering.

# _Algorithm_
Compute wpd for each granularity.The distance matrix is then computed with 20 time series as cases and three granularities as variables. Manhattan distances are computed and then hierarchical clustering is applied on them.

```{r}
t <- seq(0,299, 1)

g1 = rep(c(0, 1), length.out= length(t))
g1_dist_design1 <- rep(c(0, 0), length.out= length(t))
g1_dist_design2 <- rep(c(0, 2), length.out= length(t))
g1_dist_design3 <- rep(c(1, 4), length.out= length(t))
g1_dist_design4 <- rep(c(2, 4), length.out= length(t))

g2 = rep(c(0, 1, 2), length.out= length(t))
g2_dist_design1 <- rep(c(0, 0, 0), length.out= length(t))
g2_dist_design2 <- rep(c(1, 5, 7), length.out= length(t))
g2_dist_design3 <- rep(c(1, 2, 3), length.out= length(t))
g2_dist_design4 <- rep(c(0, 2, 4), length.out= length(t))

g3 = rep(c(0, 1, 2, 3), length.out= length(t))
g3_dist_design1 <- rep(c(0, 0, 0, 0), length.out= length(t))
g3_dist_design2 <- rep(c(0, 0, 0, 0), length.out= length(t))
g3_dist_design3 <- rep(c(0, 0, 0, 0), length.out= length(t))
g3_dist_design4 <- rep(c(0, 0, 0, 0), length.out= length(t))

time_series <- bind_rows(
  bind_cols(t, g1, g2, g3, g1_dist_design1, g2_dist_design1, g3_dist_design1),
  bind_cols(t, g1, g2, g3,  g1_dist_design2, g2_dist_design2, g3_dist_design2),
  bind_cols(t,  g1, g2, g3, g1_dist_design3, g2_dist_design3, g3_dist_design3),
  bind_cols(t,  g1, g2, g3, g1_dist_design4, g2_dist_design4, g3_dist_design4), .id = "design") %>%
  set_names(c("design", "t", "g1", "g2", "g3", "g1_dist", "g2_dist", "g3_dist")) %>% 
  mutate(sum_dist = g1_dist + g2_dist + g3_dist)

 time_series %>% group_by(design) %>% summarize(s = unique(sum_dist))
```

```{r}

generate_design <-  function (t = NULL, 
                           mu = NULL){
t <- seq(0, t, 1)
g1 <- t %%2
g2 <- t %%3
g3 <- t %%4

# design2 (change across g2 only)
mu = mu
len_mu <- length(mu)
len_obs <-  length(t)/length(mu)
dist <- dist_normal(mu = mu, sigma = 1)
numbers <- generate(dist, times = floor(len_obs))

timeseries <- numbers%>% 
  unlist() %>%
  matrix(ncol = len_mu, nrow = len_obs, byrow = FALSE) %>%
  t() %>% 
  matrix(ncol=1, byrow = TRUE)

timeseries <- as_tibble(timeseries) %>% 
  set_names("value") %>%
  bind_cols(t %>% as_tibble()) %>% 
  set_names(c("value", "t"))

 data <- tibble::tibble(t, g1, g2, g3) %>% left_join(timeseries)
 data
}

set.seed(12345)
data_design_1 <- generate_design(t = 299, mu = 0)
data_design_2 <- generate_design(t = 299, mu = c(1, 7, 7, 3, 5, 9))
data_design_3 <- generate_design(t = 299, mu = c(2, 6, 4, 5, 3, 7))
data_design_4 <- generate_design(t = 299, mu = c(2, 6, 6, 4, 4, 8))
```

<!-- # Design1 -->
```{r design-null, eval = FALSE}
p1 <- ggplot(data_design_1, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw() 
p2 <- ggplot(data_design_1, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw()
p3 <- ggplot(data_design_1, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() 
p4 <- ggplot(data_design_1, aes(x = as.factor(g3), y = value)) + geom_boxplot() +
  xlab("g3") + theme_bw()

(p1 /(p2 + p3 + p4)) 
```

<!-- # Design2 -->
```{r design1, eval = FALSE}
p1 <- ggplot(data_design_2, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw()
p2 <- ggplot(data_design_2, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw()
p3 <- ggplot(data_design_2, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw()
p4 <- ggplot(data_design_2, aes(x = as.factor(g3), y = value)) + geom_boxplot() +
  xlab("g3") + theme_bw()

(p1 /(p2 + p3 + p4)) 
```

<!-- # Design3 -->

```{r design2, eval = FALSE}
p1 <- ggplot(data_design_3, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw()
p2 <- ggplot(data_design_3, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw()
p3 <- ggplot(data_design_3, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw()
p4 <- ggplot(data_design_3, aes(x = as.factor(g3), y = value)) + geom_boxplot() +
  xlab("g3") + theme_bw()

(p1 /(p2 + p3 + p4)) 
```

<!-- # Design4 -->

```{r design3, eval = FALSE}
p1 <- ggplot(data_design_4, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw()
p2 <- ggplot(data_design_4, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw()
p3 <- ggplot(data_design_4, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw()
p4 <- ggplot(data_design_4, aes(x = as.factor(g3), y = value)) + geom_boxplot() +
  xlab("g3") + theme_bw()

(p1 /(p2 + p3 + p4))

```

# Raw plots
```{r all-design}
bind_data <- bind_rows(data_design_1, data_design_2, data_design_3, data_design_4, .id = "design") %>% mutate(design = paste("design", design, sep = "-"))

p1 <- ggplot(bind_data, aes(x = t, y = value)) + geom_line() + xlab("time")  + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0))) 
p2 <- ggplot(bind_data, aes(x = as.factor(g1), y = value)) + geom_boxplot(width=0.5) + xlab("g1") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+stat_summary(fun.y = median, geom="line", group= 1, color= "blue", size = 0.5)
p3 <- ggplot(bind_data, aes(x = as.factor(g2), y = value)) + geom_boxplot(width=0.5) + xlab("g2") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+stat_summary(fun.y = median, geom="line", group= 1, color= "blue", size = 0.5)

p4 <- ggplot(bind_data, aes(x = as.factor(g3), y = value)) + geom_boxplot(width=0.5) +
  xlab("g3") + theme_bw() + facet_wrap(~design, ncol = 1, scales = "free_y") + theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0))) +stat_summary(fun.y = median, geom="line", group= 1, color= "blue", size = 0.5)

p1
```

# Designs (Distribution of simulated data across different granularities)

```{r, out.height="100%"}
(p2+p3+p4)

```

```{r}
seed_len = seq(12345, 12399, 1)
bind_data_iter <- map(seq_len(length(seed_len)), function(x){
set.seed(seed_len[x])
data_design_1 <- generate_design(t = 299, mu = 0)
data_design_2 <- generate_design(t = 299, mu = c(1, 7, 7, 3, 5, 9))
data_design_3 <- generate_design(t = 299, mu = c(2, 6, 4, 5, 3, 7))
data_design_4 <- generate_design(t = 299, mu = c(2, 6, 6, 4, 4, 8))
bind_data <- bind_rows(data_design_1, data_design_2, data_design_3, data_design_4, .id = "design") %>% mutate(design = paste("design", design, sep = "-"))
}) %>% bind_rows(.id = "seed_id")

```

<!-- Cross-check -->
<!-- data = bind_data_iter %>% filter(design == "design-2", seed_id == 5) -->
<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() + facet_wrap(~design, ncol = 1) -->
<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() + facet_wrap(~design, ncol = 1) -->
<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
<!-- xlab("g3") + theme_bw() + facet_wrap(~design, ncol = 1) -->
<!-- (p2 + p3 + p4) -->

<!-- ```{r} -->
<!-- p2 <- ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot() + xlab("g1") + theme_bw() -->
<!-- p3 <- ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot() + xlab("g2") + theme_bw() -->
<!-- p4 <- ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot() + -->
<!-- xlab("g3") + theme_bw()  -->
<!-- (p2 + p3 + p4) -->
<!-- ``` -->

```{r}
harmony_tbl <- tibble(facet_variable = NA, x_variable = c("g1", "g2", "g3"),
                      x_levels = c(2, 3, 4), facet_levels = NA)
```

```{r wpd-table}
data <- bind_data_iter %>% mutate(data_id = paste(design, seed_id, sep = "-"))
bind_data_iter <- tsibble::as_tsibble(data, index = t, key = "data_id")

data_wpd <- map(seq_len(length(unique(bind_data_iter$data_id))),
    function(y){
      z = unique(bind_data_iter$data_id)
      data_sub <- bind_data_iter %>% filter(data_id==z[y])
        elec_select_harmony = hakear::select_harmonies(data_sub,
                               harmony_tbl = harmony_tbl,
                               response = value,
                               nperm = 200
  ) %>% mutate(data_id = unique(data_sub$data_id)) %>% unnest()
    }) %>% bind_rows()

#write_csv(data_wpd, "data_wpd.rds")
```

# Table of wpd across designs and wpd

```{r data-wide}

data_wide <- data_wpd %>% select(x_variable, wpd, data_id) %>% 
  pivot_wider(names_from = x_variable, values_from = wpd)
data_wide %>% kable()

ggpairs_nocolor <- ggpairs(data_wide, columns = 2:4)

rownames(data_wide) = data_wide$data_id

hc = stats::hclust(dist(data_wide[-1], method = "manhattan"),method="complete")

library(fpc)
koptimal = fpc::nselectboot(dist(data_wide[-1], method="manhattan"),
                        B = 500,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:8)

koptimal = koptimal$kopt
```

# Clustering of designs

```{r}
  groups <- cutree(hc, k=4) %>% as_tibble() %>% 
mutate(data_id = data_wide$data_id) %>% rename("group" = "value")

  groups %>% kable()

data_wide_group <- data_wide %>% left_join(groups)  %>% 
  mutate(group_id = as.factor(group))

parcoord <- GGally::ggparcoord(data_wide_group,
                   columns = 2:ncol(data_wide),
                   groupColumn = "group_id",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.8,
                   scale = "globalminmax", 
                   order = "anyClass"
) + ggplot2::theme_bw() +
  scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")
parcoord
ggpairs_nocolor
  # ggpairs_color <- ggscatmat(data_wide, columns=2:4, color = "group") +
  # scale_colour_brewer(palette="Set2")
```

# PCA based clustering

```{r, eval = FALSE}
pc <- prcomp(data_wide_group[c(2, 3, 4)], center = F, scale = F)

# First two principal componentsafc
comp <- data.frame(pc$x[,1:2])

#plot(comp, pch=16, col=rgb(0,0,0,0.5))

hc = stats::hclust(dist(comp))
groups <- cutree(hc, k=koptimal) %>% as_tibble() %>% bind_cols(comp) %>% 
  mutate(value = as.character(value))
library(RColorBrewer)
library(scales)
groups %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = value)) +
  scale_color_brewer(palette = "Dark2") + theme_bw()
```

```{r}
data_plot <- left_join(data_wide_group,  bind_data_iter %>% distinct(design, data_id)) %>%
  mutate(pred_group = case_when(
    group == 1 ~"design-1",
    group == 2 ~"design-2",
    group == 3 ~"design-3",
    group == 4 ~"design-4"
  )) %>% 
  mutate(pred_group = as.factor(pred_group),
         design = as.factor(design)
  )

xtab <- confusionMatrix(data_plot$design, data_plot$pred_group)

data_plot %>%  ggplot(aes(x=g1, y=g2)) + geom_point(aes(colour = design)) +
  scale_color_brewer(palette = "Dark2")

left_join(data_wide_group,  bind_data_iter %>% distinct(design, data_id)) %>%  ggplot(aes(x=g1, y=g3)) + geom_point(aes(colour = design)) +
  scale_color_brewer(palette = "Dark2")

left_join(data_wide_group,  bind_data_iter %>% distinct(design, data_id)) %>%  ggplot(aes(x=g2, y=g3)) + geom_point(aes(colour = design)) +
  scale_color_brewer(palette = "Dark2")

```