### 2021-08-18 supplementary/algorithm3/real-data/handpick-18.Rmd
simulated and real data results update for algorithm 3

---
title: "Hand picking similar behaving group of customers to check clustering results"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE, message=FALSE, warning=FALSE)
library(readr)
library(tidyverse)
library(gravitas)
library(tsibble)
library(parallel)
library(magrittr)
library(GGally)
library(knitr)
library(patchwork)
library(caret)
```

```{r read-data}
data <- read_rds("../../../data/elec_nogap_2013_100.rds")
```

```{r hand-pick}

# data_pick <- data %>% 
#   filter(customer_id %in% 
#                    c(10538785, 8530300, 8717912, 8171227, 10108102, 10767756, 10895401, 10704368, 9397229, 8281474))

data_pick_onepeak <- 
  data %>% 
  filter(customer_id %in% c(8485375,10690820,10859082,8486921, 10208124,8198401))%>% mutate(design = 1)

data_pick_threepeak <- 
  data %>% 
  filter(customer_id %in% c(11013154, 10046512, 8637773, 8952846, 8196501, 8627007))%>% mutate(design = 3)

data_pick_twopeak <- 
  data %>% 
  filter(customer_id %in% c(8181071, 8955520, 8640839, 10469265, 10538785, 10895401))%>% mutate(design = 2)

data_pick <- bind_rows(data_pick_onepeak, data_pick_twopeak, data_pick_threepeak)

elec_harmony_all <- read_rds("../../../data/elec_harmony-100-nogap-new.rds")%>%
  filter(customer_id %in% data_pick$customer_id)

elec_harmony_hod <- elec_harmony_all %>% filter(x_variable=="hour_day") %>% 
  mutate(wpd = round(wpd, 2))

group2_data <- data_pick %>% 
  create_gran("hour_day") %>% 
  as_tibble() %>% 
  group_by(design, hour_day, customer_id) %>% 
  summarise(median_kwh = median(general_supply_kwh))

# 
# group2_data%>% 
#   ungroup() %>% 
#   filter(design==1) %>% 
#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
#   geom_line(size = 0.5) +
#   facet_wrap(~customer_id, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
# 
# 
# group2_data%>% 
#   ungroup() %>% 
#   filter(design==2) %>% 
#   #filter(customer_id %in% c(8148781, 10703878)) %>% 
#   ggplot(aes(x = as.integer(hour_day), y = median_kwh)) +
#   geom_line(size = 0.5) +
#   facet_wrap(~customer_id, scales = "free_y") + theme_minimal()+ theme(strip.text.y = element_text(size = 5))
```

```{r characterize, fig.cap= "Median (black) and quartile deviation (blue region) of hourly demand drawn for few customers showing similar behaviors. Roughly speaking, Design 1 has one evening peak, Design 2 has two peaks and Design 3 has three peaks in a day. Each of design 1, 2 and 3 have six similar behaving customers resulting to 18 time series. We want our clustering results to group each of the designs together."}
quantile_prob_val = c(0.25, 0.5, 0.75)

sm_hod <- gravitas::create_gran(data_pick, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(design, customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

# groups_ref <- unique(data_pick$design) %>% as_tibble %>% 
#               mutate(customer_serial_id = sm_hod_list$customer_id)

# sm_hod_list_cat <-  sm_hod_list %>% left_join(groups_ref, by = c("customer_id" = "customer_serial_id")) %>% select(group, everything())

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, 2)]))
nrow_sm <- 1:18

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, 2)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, 2)] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>% 
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id, design))

# 
# %>% 
#   left_join(categories_ref_cat) %>% 
#   rename("category" = "value") %>% 
#   select(-categories_serial_id) %>% 
#   mutate(category = as.numeric(category))

data_heatmap %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`), size = 0.7) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)")
```
# quantile_prob_val = c(0.5, 0.75)

## only hod

```{r,echo = TRUE}
quantile_prob_val = c(0.5, 0.75)
#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
library(gracsr)
v2 <- suppressWarnings(
  scaled_dist_gran(data_pick, "hour_day", 
                   response = "general_supply_kwh",
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
v3 <- suppressWarnings(
  scaled_dist_gran(data_pick, "day_month",
                   response = "general_supply_kwh", 
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")

data_dist <- v3 %>% 
  left_join(v2) %>%
  mutate(dist = dist_hod) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")
```

```{r}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="complete")

groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())
#groups %>% filter(group==1)
#groups %>% filter(group==2)
#groups %>% filter(group==3)

data_heatmap_new <- data_heatmap %>% left_join(groups)

data_heatmap_new$group = as.factor(data_heatmap_new$group)

data_heatmap_new %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`, color = group), size = 0.7) +
  facet_wrap(group~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2")

```

## hod + dom

```{r,echo = TRUE}
#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
library(gracsr)
v2 <- suppressWarnings(
  scaled_dist_gran(data_pick, "hour_day", 
                   response = "general_supply_kwh",
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
v3 <- suppressWarnings(
  scaled_dist_gran(data_pick, "day_month",
                   response = "general_supply_kwh", 
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")

data_dist <- v3 %>% 
  left_join(v2) %>%
  mutate(dist = dist_hod + dist_dom) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")
```

```{r}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="complete")

groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())
#groups %>% filter(group==1)
#groups %>% filter(group==2)
#groups %>% filter(group==3)

data_heatmap_new <- data_heatmap %>% left_join(groups)

data_heatmap_new$group = as.factor(data_heatmap_new$group)

data_heatmap_new %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`, color = group), size = 0.7) +
  facet_wrap(group~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2")

```

# quantile_prob_val = seq(0.1, 0.9, 0.1)

## only hod

```{r,echo = TRUE}
quantile_prob_val = seq(0.1, 0.9, 0.1)
#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
library(gracsr)
v2 <- suppressWarnings(
  scaled_dist_gran(data_pick, "hour_day", 
                   response = "general_supply_kwh",
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
v3 <- suppressWarnings(
  scaled_dist_gran(data_pick, "day_month",
                   response = "general_supply_kwh", 
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")

data_dist <- v3 %>% 
  left_join(v2) %>%
  mutate(dist = dist_hod) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")
```

```{r}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="complete")

groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())
#groups %>% filter(group==1)
#groups %>% filter(group==2)
#groups %>% filter(group==3)

data_heatmap_new <- data_heatmap %>% left_join(groups)

data_heatmap_new$group = as.factor(data_heatmap_new$group)

data_heatmap_new %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`, color = group), size = 0.7) +
  facet_wrap(group~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2")

```

## hod + dom

```{r,echo = TRUE}
#data_pick <- data_pick %>% filter(!(customer_id %in% c(8485375, 8952846)))
library(gracsr)
v2 <- suppressWarnings(
  scaled_dist_gran(data_pick, "hour_day", 
                   response = "general_supply_kwh",
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
v3 <- suppressWarnings(
  scaled_dist_gran(data_pick, "day_month",
                   response = "general_supply_kwh", 
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")

data_dist <- v3 %>% 
  left_join(v2) %>%
  mutate(dist = dist_hod + dist_dom) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")
```

```{r}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="complete")

groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())
#groups %>% filter(group==1)
#groups %>% filter(group==2)
#groups %>% filter(group==3)

data_heatmap_new <- data_heatmap %>% left_join(groups)

data_heatmap_new$group = as.factor(data_heatmap_new$group)

data_heatmap_new %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`, color = group), size = 0.7) +
  facet_wrap(group~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2")

```

## remove the outlier and run again

```{r,echo = TRUE}
quantile_prob_val = c(0.5, 0.75)
data_pick <- data_pick %>% filter(!(customer_id %in% 8952846))
library(gracsr)
v2 <- suppressWarnings(
  scaled_dist_gran(data_pick, "hour_day", 
                   response = "general_supply_kwh",
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_hod" = "dist")
v3 <- suppressWarnings(
  scaled_dist_gran(data_pick, "day_month",
                   response = "general_supply_kwh", 
                   quantile_prob_val = quantile_prob_val)) %>% rename("dist_dom" = "dist")

data_dist <- v3 %>% 
  left_join(v2) %>%
  mutate(dist = dist_hod + dist_dom) %>% 
    pivot_wider(-c(3, 4), 
                names_from = customer_to,
              values_from = dist) %>%
  rename("customer_id" = "customer_from")
```

```{r}
d = stats::as.dist(data_dist[,-1])
hc = stats::hclust(d,method="complete")

groups <- tibble(group = cutree(hc, k=3), customer_id = data_dist$customer_id)
groups %>% group_by(group) %>% summarise(n=n())
#groups %>% filter(group==1)
#groups %>% filter(group==2)
#groups %>% filter(group==3)

data_heatmap_new <- data_heatmap %>% filter(customer_id!=8952846) %>% left_join(groups)

data_heatmap_new$group = as.factor(data_heatmap_new$group)

data_heatmap_new %>% 
  ggplot(aes(x = as.integer(category))) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`, color = group), size = 0.7) +
  facet_wrap(group~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") +
  scale_color_brewer(palette = "Dark2")

```