### 2021-08-18 supplementary/algorithm2/3cat-same-lev.Rmd

---
title: "Simulation for wpd algorithm"
author: "Sayani Gupta"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(distributional)
```

```{r}
T <- seq(0, 500, 1)
g1 <- T %%2
g2 <- T %%3

g3 <- T %%4
data <- tibble::tibble(T, g1, g2, g3)

generate_dist <- function(x){
  if(x==0) 
    sim_dist <- distributional::dist_normal(0,1)
  if(x==1)
    sim_dist <- distributional::dist_normal(2,1)
  if(x==2)
    sim_dist <- distributional::dist_normal(4,1)
  if(x==3)
    sim_dist <- distributional::dist_normal(6,1)
  return(sim_dist)
}

set.seed(12345)

y <- data %>% group_split(g1)
sim_data_g1 <- map(seq_len(length(y)), 
      function(z){
  univec =  y %>% magrittr::extract2(z,) %>% select(T, g1)
  value = univec %>% distinct(g1) %>% pull(g1)
  return_value = generate_dist(value) %>% distributional::generate(nrow(univec)) %>% unlist() %>% as_tibble() %>% set_names("value") %>% 
    bind_cols(univec)
  }) %>% bind_rows() %>% select(T, g1, value)

g2_value = generate(dist_normal(mu = 0, sigma = 1), times = nrow(sim_data_g1)) %>% unlist() %>% as_tibble() %>% set_names("g2_value")

g3_value = generate(dist_normal(mu = 2, sigma = 1), times = nrow(sim_data_g1)) %>% unlist() %>% as_tibble() %>% set_names("g3_value")

total_data <- sim_data_g1 %>% 
  arrange(T) %>% 
  left_join(data) %>% 
  rename("g1_value" = "value") %>% 
  bind_cols(g2_value, g3_value) %>% 
  select(T, g1, g2, g3, everything()) %>% 
  group_by(g1, g2, g3) %>% 
  mutate(sim_data = sum(g1_value, g2_value, g3_value))

ggplot(total_data, aes(x = as.factor(g1), y = sim_data)) + geom_boxplot()
ggplot(total_data, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot()
ggplot(total_data, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot()

```

```{r}
y <- data %>% group_split(g2)

sim_data_g2 <- map(seq_len(length(y)), 
      function(z){
  univec =  y %>% magrittr::extract2(z,) %>% select(T, g2)
  value = univec %>% distinct(g2) %>% pull(g2)
  return_value = generate_dist(value) %>% distributional::generate(nrow(univec)) %>% unlist() %>% as_tibble() %>% set_names("value") %>% 
    bind_cols(univec)
  }) %>% bind_rows() %>% select(T, g2, value)

ggplot(sim_data_g2, aes(x = as.factor(g2), y = value)) + geom_boxplot()
```

```{r}
y <- data %>% group_split(g3)

sim_data_g3 <- map(seq_len(length(y)), 
      function(z){
  univec =  y %>% magrittr::extract2(z,) %>% select(T, g3)
  value = univec %>% distinct(g3) %>% pull(g3)
  return_value = generate_dist(value) %>% distributional::generate(nrow(univec)) %>% unlist() %>% as_tibble() %>% set_names("value") %>% 
    bind_cols(univec)
  }) %>% bind_rows() %>% select(T, g3, value)

ggplot(sim_data_g3, aes(x = as.factor(g3), y = value)) + geom_boxplot()
```
# sum up the normal distribution approach

```{r}
set.seed(12345)
dist <- dist_normal(mu = c(0,4,4,8), sigma = 1)
numbers <- generate(dist, times = 100)

timeseries <- numbers%>% unlist() %>% matrix(ncol = 4, nrow = 100, byrow = FALSE) %>% t() %>% 
  matrix(ncol=1, byrow = TRUE)

timeseries <- as_tibble(timeseries) %>% set_names("value") %>% bind_cols(t = seq(0, 399, 1))
ggplot(timeseries, aes(x = t, y = value)) + geom_line()

t <- seq(0,399, 1)
g1 <- t %%2
g2 <- t %%3

g3 <- t %%4
data <- tibble::tibble(t, g1, g2, g3) %>% left_join(timeseries)

ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot()
ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot()
ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot()

```

```{r}
set.seed(12345)
dist <- dist_normal(mu = c(0,2,4), sigma = 1)
numbers <- generate(dist, times = 100)

timeseries <- numbers%>% unlist() %>% matrix(ncol = 3, nrow = 100, byrow = FALSE) %>% t() %>% 
  matrix(ncol=1, byrow = TRUE)

timeseries <- as_tibble(timeseries) %>% set_names("value") %>% bind_cols(t = seq(0, 299, 1))
ggplot(timeseries, aes(x = t, y = value)) + geom_line()

t <- seq(0,299, 1)
g1 <- t %%2
g2 <- t %%3

g3 <- t %%4
data <- tibble::tibble(t, g1, g2, g3) %>% left_join(timeseries)

ggplot(data, aes(x = as.factor(g1), y = value)) + geom_boxplot()
ggplot(data, aes(x = as.factor(g2), y = value)) + geom_boxplot()
ggplot(data, aes(x = as.factor(g3), y = value)) + geom_boxplot()
```