### 2021-08-18 supplementary/algorithm2/real-data/algo2-elec100-diffscale.Rmd

---
title: "Algorithm using wpd values for clustering 100 households"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
```

```{r load-lib}
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
#library(ggridges)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)
library(ggplot2)
library(patchwork)
library(scales)
library(GGally)
library(viridis)
```

```{r allplot, message=FALSE, warning=FALSE, echo = FALSE}

##----harmony-all
elec_harmony_all <- read_rds("../data/algo2-rawdata-wpd.rds")

select_split <- str_split(elec_harmony_all$select_harmony, " ", simplify = TRUE)[,2]

elec_sig_split <- elec_harmony_all %>%
  mutate(facet_variable = case_when(
    facet_variable == "hour_day" ~ "hod" ,
    facet_variable == "day_month" ~ "dom" ,
    facet_variable == "day_week" ~ "dow" ,
    facet_variable == "week_month" ~ "wom" ,
    facet_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  mutate(x_variable = case_when(
    x_variable == "hour_day" ~ "hod" ,
    x_variable == "day_month" ~ "dom" ,
    x_variable == "day_week" ~ "dow" ,
    x_variable == "week_month" ~ "wom" ,
    x_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  bind_cols(select_split = select_split) %>% 
  mutate(significant = case_when(
    select_split == "***" ~ "highest",
    select_split == "**" ~ "high",
    select_split == "*" ~ "medium",
    select_split == "" ~ "low"
  )) %>%mutate(comb = paste(facet_variable, x_variable, sep = "-"))

# %>% 
#   mutate(rank = case_when(
#     select_split == "***" ~ paste(rank, "***", sep = " "),
#     select_split == "**" ~  paste(rank, "**", sep = " "),
#     select_split == "*" ~  paste(rank, "*", sep = " "),
#     select_split == "" ~  paste(rank, "", sep = " ")
#   ))

elec_sig_split$significant <- 
  factor(elec_sig_split$significant, levels = c("highest", "high", "medium", "low"))

data_comb_filter = elec_sig_split %>% 
mutate(low_proxy = if_else(significant=="low", 0, 1))%>% 
select(comb, customer_id, low_proxy) %>% 
pivot_wider(names_from = customer_id, values_from = low_proxy) %>% 
mutate(score = rowMeans(across(-1))) %>% 
filter(score !=0)

data_sig_filter = elec_sig_split %>% 
filter(comb %in% data_comb_filter$comb)

data_clust = data_sig_filter %>% 
select(comb, customer_id, wpd) %>% 
pivot_wider(names_from = comb, values_from = wpd)

hc = stats::hclust(dist(data_clust[-1], method = "manhattan"),method="complete")
#plot(hc)
plot(hc, cex = 0.6)
rect.hclust(hc, k = 8, border = 2:10)

library(fpc)
koptimal = fpc::nselectboot(dist(data_clust[-1], method = "manhattan"),
                        B = 50,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:8)

koptimal = koptimal$kopt

groups <- cutree(hc, k=8) %>% 
  as_tibble() %>% 
mutate(customer_id = data_clust$customer_id)

##----cluster-characterization

data = data_clust %>% 
left_join(groups)%>% 
             rename ("group" = "value") %>% 
             ungroup() %>% 
             select(group, customer_id, everything()) %>% 
             pivot_longer(-c(1,2), names_to = "comb", values_to = "wpd")

data_pcp <- data %>% pivot_wider(
              names_from = comb, 
              values_from = wpd) %>% 
  mutate(group = as.character(group)) 
```

```{r}
ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
ggpairs_nocolor

groupge2 <- data %>% 
  group_by(group) %>% 
  summarise(n = length(unique(customer_id))) %>% 
  filter(n>2)

data_pcp_ge2 <- data_pcp %>% filter(group %in% groupge2$group)

ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)

ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:7, color = "group") +
  scale_colour_brewer(palette="Set2")
ggpairs_color

# 
# data_order <- data %>% 
#   group_by(comb) %>% 
#   summarise(m = mean(wpd)) %>% 
#   arrange(desc(m))
#   
# data_median = data %>% 
#   group_by(group, comb) %>% 
#   summarize(wpd_med = median(wpd))
# 
# data_pcp <- data_median %>% 
# mutate(group = as.character(group)) %>% 
#   left_join(data_order) %>% 
#   arrange(desc(m)) %>% 
#   pivot_wider(-4, 
#               names_from = comb, 
#               values_from = wpd_med)

data_pcp <- data_pcp %>% ungroup()%>% filter(group %in% groupge2$group)
parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 3:ncol(data_pcp),
                   groupColumn = "group",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.3,
                   order = "allClass",
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")

c(6, 7, 3, 4, 5)
# 
# ggparcoord(data_pcp, scale = "globalminmax", columns=3:ncol(data_pcp),groupColumn = "group", alphaLines = 0.8) + geom_boxplot(aes_string(group = "variable"), width = 0.1, alpha = 0.1,
#                  outlier.color = NA) +  scale_color_brewer(palette = "Dark2") 
#+
 # geom_point(data = data, aes(x=comb, y = wpd))

##----read-orig-data

elec <- read_rds("../data/elec_nogap_2013_100.rds") %>% 
  left_join(groups)%>% 
             rename ("group" = "value")

# plot(hc, cex = 0.6)
# rect.hclust(hc, k = 4, border = 2:7)

#elec_zoom
parcoord

elec_scaled <- elec %>%
    group_by(customer_id) %>%
    #dplyr::mutate(general_supply_kwh = scale(general_supply_kwh)) %>%
    ungroup()

sm_hod_dow <- elec_scaled %>% 
gravitas::create_gran("hour_day") %>% 
  gravitas::create_gran("day_week")

sm_hod_list <- sm_hod_dow %>% 
  as_tibble() %>% 
  arrange(group, day_week, hour_day) %>% 
  mutate(category = paste(hour_day, day_week, sep = "-")) %>% 
   select(group, category, general_supply_kwh) %>% 
  pivot_wider(names_from = category,
              values_from = general_supply_kwh)

elec %>% as_tibble() %>% group_by(group) %>% summarise(n = length(unique(customer_id)))

quantile_prob_val = c(0.25, 0.5, 0.75)

p1 <- sm_hod_dow %>% 
  filter(group==1) %>% 
  gravitas::prob_plot(
     "day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 1") + theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p2 <- sm_hod_dow %>% 
  filter(group==2 ) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10()+
    ggtitle("Group 2")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p3 <- sm_hod_dow %>% 
  filter(group==3) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 3" )+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p4 <- sm_hod_dow %>% 
  filter(group==4) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 4")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p5 <- sm_hod_dow %>% 
  filter(group==5) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 5")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p6 <- sm_hod_dow %>% 
  filter(group==6) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 6")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p7 <- sm_hod_dow %>% 
  filter(group==7) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 7")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines"))+ ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

p8 <- sm_hod_dow %>% 
  filter(group==8) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 8")+ theme_bw()  +
  theme(legend.position = "none",
        strip.text = element_text(size = 10, margin = margin())) + 
  theme(panel.spacing =unit(0, "lines")) + ylab("")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, size = 5))

(p1 + p2)/ (p3 + p4)
(p5 + p6)/(p7 +p8)

```