### 2021-08-18 supplementary/algorithm2/real-data/algo2-elec8.Rmd

---
title: "Algorithm using wpd values for clustering 8 households"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
```

```{r load-lib}
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
#library(ggridges)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)
library(ggplot2)
library(patchwork)
library(scales)
library(GGally)
library(viridis)
```

```{r, message=FALSE, warning=FALSE, echo = FALSE}

##----harmony-all
elec_harmony_all <- read_rds("../../paper-hakear/paper/data/elec_harmony_all.rds")

select_split <- str_split(elec_harmony_all$select_harmony, " ", simplify = TRUE)[,2]

elec_sig_split <- elec_harmony_all %>% 
  bind_cols(select_split = select_split) %>% 
  mutate(significant = case_when(
    select_split == "***" ~ "highest",
    select_split == "**" ~ "high",
    select_split == "*" ~ "medium",
    select_split == "" ~ "low"
  )) %>% 
  mutate(rank = case_when(
    select_split == "***" ~ paste(rank, "***", sep = " "),
    select_split == "**" ~  paste(rank, "**", sep = " "),
    select_split == "*" ~  paste(rank, "*", sep = " "),
    select_split == "" ~  paste(rank, "", sep = " ")
  ))%>%mutate(comb = paste(facet_variable, x_variable, sep = "-"))

elec_sig_split$significant <- 
  factor(elec_sig_split$significant, levels = c("highest", "high", "medium", "low"))

data_comb_filter = elec_sig_split %>% 
mutate(low_proxy = if_else(significant=="low", 0, 1))%>% 
select(comb, id, low_proxy) %>% 
pivot_wider(names_from = id, values_from = low_proxy) %>% 
mutate(score = rowMeans(across(`id 1`:`id 8`))) %>% 
filter(score !=0)

data_sig_filter = elec_sig_split %>% 
filter(comb %in% data_comb_filter$comb)

data_clust = data_sig_filter %>% 
select(comb, id, wpd) %>% 
pivot_wider(names_from = comb, values_from = wpd)

hc = stats::hclust(dist(data_clust[-1], method="manhattan"),method="complete")
#plot(hc)

library(fpc)
koptimal = fpc::nselectboot(dist(data_clust[-1], method="manhattan"),
                        B = 50,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:5)

koptimal = koptimal$kopt
koptimal

groups <- cutree(hc, k=4) %>% as_tibble() %>% 
mutate(id = data_clust$id)

groups %>% 
  dplyr::group_by(value) %>%  
  summarise(n = n())

data_wide <- data_clust %>% left_join(groups) %>%
  rename("group" = "value") %>% 
  mutate(group = as.factor(group)) %>% ungroup()

##----cluster-characterization

parcoord <- GGally::ggparcoord(data_wide,
                   columns = 2:(ncol(data_wide)-1),
                   groupColumn = "group",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.8,
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")

# data = data_clust %>% 
# left_join(groups)%>% 
#              rename ("group" = "value") %>% 
#              ungroup() %>% 
#              select(group, everything(), -id) %>% 
#              pivot_longer(-1, names_to = "comb", values_to = "wpd")
# 
# data_order <- data %>% 
#   group_by(comb) %>% 
#   summarise(m = mean(wpd)) %>% 
#   arrange(desc(m))
#   
# data_median = data %>% 
#   group_by(group, comb) %>% 
#   summarize(wpd_med = median(wpd))
# 
# data_pcp <- data_median %>% 
# mutate(group = as.character(group)) %>% 
#   left_join(data_order) %>% 
#   arrange(desc(m)) %>% 
#   pivot_wider(-4, 
#               names_from = comb, 
#               values_from = wpd_med) 
#              
# parcoord <- GGally::ggparcoord(data_pcp,
#                    columns = 2:ncol(data_pcp),
#                    groupColumn = 1,
#                    showPoints = TRUE, 
#                    title = "(c)",
#                    alphaLines = 0.8,
#                    scale = "globalminmax"
# ) + ggplot2::theme_bw() +
#   scale_color_viridis(discrete=TRUE) + 
#   ggplot2::theme(
#     plot.title = ggplot2::element_text(size=10)
#   )+
#   ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
#   theme(legend.position = "bottom") +
#   xlab("") +
#   ylab("wpd")
```

```{r}
##----read-orig-data

elec <- read_rds(here("../paper-hakear/paper/data/elec_all-8.rds")) %>% 
  dplyr::filter(date(reading_datetime) >= ymd("20190701"), date(reading_datetime) < ymd("20191231"), meter_id==1) %>% 
  select(-meter_id) %>% 
  rename("id" = "household_id",
         "date_time" = "reading_datetime") %>% 
  mutate(date = date(date_time)) %>% 
  mutate(id = paste("id", id, sep = " ")) %>% 
  left_join(groups)%>% 
             rename ("group" = "value")

elec_zoom <-  elec %>%
  as_tibble() %>% 
  dplyr::filter(date(date_time) > as.Date("2019-08-01") & date(date_time) < (as.Date("2019-09-30"))) %>%
  ggplot(aes(x=date_time, y = kwh)) +
  #geom_point(size = 0.1, colour = "black", alpha = 0.3) +
  geom_line(size = 0.1, colour = "blue") +
  facet_wrap(~id, 
             scales = "free_y",
             ncol = 1,
             strip.position =  "right") + 
  xlab("Time [30m]") + 
  theme_grey() + 
  ylab("Energy demand (in Kwh)") + ggtitle("(b)") +
  theme(panel.grid.major.x = element_blank()) +
  scale_x_datetime("Date", date_labels = "%b %d",
                   breaks = "1 week",
                   date_minor_breaks = "1 day")  + theme_bw() +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_line(colour = "#D3D3D3")) +
  theme(
    strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))

plot(hc, cex = 0.6)
rect.hclust(hc, k = 4, border = 2:7)

elec_zoom
parcoord

ggpairs_nocolor <- ggpairs(data_wide, columns = 2:(ncol(data_wide)-1))
ggpairs_nocolor

sm_hod_dow <- elec %>% 
gravitas::create_gran("hour_day") %>% 
  gravitas::create_gran("day_week") %>% 
    dplyr::mutate(kwh = scale(kwh))

sm_hod_list <- sm_hod_dow %>% 
  as_tibble() %>% 
  arrange(group, day_week, hour_day) %>% 
  mutate(category = paste(hour_day, day_week, sep = "-")) %>% 
   select(group, category, kwh) %>% 
  pivot_wider(names_from = category,
              values_from = kwh)

quantile_prob_val = c(0.25, 0.5, 0.75)

  sm_hod_dow %>% 
  filter(group==1) %>% 
  gravitas::prob_plot(
     "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 1")

  sm_hod_dow %>% 
  filter(group==2) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10()+
    ggtitle("Group 2")

  sm_hod_dow %>% 
  filter(group==3) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 3")

  sm_hod_dow %>% 
  filter(group==4) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 4")
```
<!-- ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)] -->
<!-- nrow_sm <- seq_len(nrow(sm_hod_list)) -->

<!-- sm_hod_quantiles <- map(nrow_sm, function(x){ -->
<!--   map(ncol_sm, function(y){ -->
<!--    cell <- sm_hod_list[-1] %>%  -->
<!--    magrittr::extract(x, y) %>% unlist() -->
<!--    quantile(cell, prob = quantile_prob_val) -->
<!-- })  %>% bind_rows(.id = "categories_serial_id") -->
<!-- }) %>% bind_rows(.id = "group_id") -->

<!-- sm_hod_quantiles_cat <- map(nrow_sm, function(x){ -->
<!--   map(ncol_sm, function(y){ -->
<!--    cell <- sm_hod_list%>%  -->
<!--      dplyr::filter(group == x) %>%  -->
<!--      select(-c(1)) %>%  -->
<!--      magrittr::extract(y) %>%  -->
<!--      unlist() -->
<!--    quantile(cell, prob = quantile_prob_val) -->
<!-- })  %>% bind_rows(.id = "categories_serial_id") -->
<!-- }) %>% bind_rows(.id = "group_id") %>%  -->
<!--   mutate(group_id = as.integer(group_id), -->
<!--          categories_serial_id = as.integer(categories_serial_id)) -->

<!-- categories_ref_cat <- names(sm_hod_list)[-c(1)] %>% as_tibble() %>%  -->
<!--   mutate(categories_serial_id = as.character(row_number())) -->

<!-- data_heatmap <- sm_hod_quantiles %>%  -->
<!--   pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>%  -->
<!--   left_join(categories_ref_cat) %>%  -->
<!--   rename("category" = "value") %>%  -->
<!--   select(-categories_serial_id) -->

<!-- data_heatmap %>%  -->
<!--   pivot_wider( names_from = quantile_prob,  -->
<!--                values_from = quantile_val) %>%  -->
<!--   ggplot(aes(x = category)) +  geom_ribbon(aes(ymin = `25%`,  -->
<!--                   ymax = `75%`), fill = "lightblue") + -->
<!--   geom_line(aes(y = `50%`)) + -->
<!--   facet_wrap(~group_id, ncol = 1, scales = "free_y", labeller = "label_both") + -->
<!--   theme_bw() + -->
<!--     theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) -->

```