### 2021-09-22 supplementary/distance-computation/real-data/algo1-cust600-gracsr.Rmd

---
title: "Algorithm using wpd values for clustering 600 households"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
```

```{r load-lib}
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
#library(ggridges)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)
library(ggplot2)
library(patchwork)
library(scales)
library(GGally)
library(viridis)
library(gracsr)
library(tsibble)
library(gravitas)
```

# distance using js distance between categories (robust scaling)

```{r js-based}
elec <- read_rds("../../../data/elec_nogap_2013_600.rds") %>% select(customer_id, reading_datetime, general_supply_kwh)
# 
# dist_mat_moy$customer_id[which(is.na(dist_mat_moy$general_supply_kwh))] %>% unique()
# c(8258490, 8662370, 10880968)

# throw away customers for which IQR is zero

elec_iqr0 <- elec %>% as_tibble() %>% group_by(customer_id) %>%
  summarise(q2 = stats::median(general_supply_kwh, na.rm = TRUE),
iqr = stats::IQR(general_supply_kwh, na.rm = TRUE)) %>% filter(iqr==0| q2==0)

dist_mat_moy <- elec %>%
  filter(!(customer_id %in% 
             elec_iqr0$customer_id)) %>% 
  scale_gran(method = "robust", response = "general_supply_kwh") %>%
  dist_gran(gran1 = "month_year", response= "general_supply_kwh")

dist_mat_hod <- elec %>%
  scale_gran(method = "robust", response = "general_supply_kwh") %>%
  dist_gran(gran1 = "hour_day", response= "general_supply_kwh")

dist_mat_wndwday <- elec %>%
  scale_gran(method = "robust", response = "general_supply_kwh") %>%
  dist_gran(gran1 = "wknd_wday", response= "general_supply_kwh")

dist_mat <- dist_mat_moy + dist_mat_hod + dist_mat_wndwday

groups = dist_mat %>% 
  clust_gran(kopt = 20)
```