### 2021-09-22 supplementary/distance-computation/real-data/algo2-cust600-gracsr.Rmd

---
title: "Algorithm using wpd values for clustering 600 households"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
```

```{r load-lib}
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
#library(ggridges)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)
library(ggplot2)
library(patchwork)
library(scales)
library(GGally)
library(viridis)
library(gracsr)
library(tsibble)
library(gravitas)
```

```{r allplot, message=FALSE, warning=FALSE, echo = FALSE}

##----harmony-all
elec_harmony_all <- read_rds("../../../data/algo2-cust600-wpd-rawdata.rds")

cust_by_wpd <- elec_harmony_all %>% 
  select(customer_id, x_variable, wpd) %>% 
  pivot_wider(
              names_from = x_variable, 
              values_from = wpd)

# Outliers

library(OutliersO3)

data <- cust_by_wpd[2:4]

O3x <- O3prep(data, method="HDo", tols=0.01, boxplotLimits=c(3, 6, 10))

O3x1 <- O3plotT(O3x)
library(gridExtra)
grid.arrange(O3x1$gO3, O3x1$gpcp, ncol=1)

cust_by_wpd <- cust_by_wpd %>% mutate(Case= row_number()) %>% 
  filter(!(Case %in% O3x1$outsTable$Case))

# O3m <- O3prep(data, method=c("HDo", "PCS"))
# O3m1 <- O3plotM(O3m)
# grid.arrange(O3m1$gO3, O3m1$gpcp, ncol=1)

library(stray)
output <- find_HDoutliers(data, knnsearchtype = "brute", alpha = 0.1)
display_HDoutliers(data, out = output)

# make a distance matrix from here and perform the algorithm

mydist <- column_to_rownames(cust_by_wpd, var = "customer_id")%>%
    dist()

groups <- mydist%>% 
  gracsr::clust_gran(kopt=10) %>% 
  bind_cols(customer_id = cust_by_wpd$customer_id)

ngroup <- groups %>% group_by(group) %>% count()

cust_by_wpd_join_groups <- cust_by_wpd %>% left_join(groups)

cust_by_wpd_join_groups_long <- cust_by_wpd_join_groups %>%
  pivot_longer(-c(1,5, 6),names_to = "gran",
               values_to = "wpd") 
```

```{r}
cust_by_wpd_join_groups_long%>% 
  ggplot() +
  geom_boxplot(aes(x=gran, y = wpd)) +
  facet_wrap(~group) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

elec <- read_rds("../../../data/elec_nogap_2013_600.rds") %>% select(customer_id, reading_datetime, general_supply_kwh)

characterise_plot_data <- elec %>% 
  create_gran("month_year") %>% 
  create_gran("hour_day") %>% 
  create_gran("wknd_wday") %>% 
  as_tibble() %>% 
 select(customer_id,
        month_year,
        hour_day,
        wknd_wday, general_supply_kwh) %>% 
left_join(groups, by = "customer_id")

# characterise_plot_data_long <- characterise_plot_data%>% 
#   pivot_longer(c(-1, -5, -6), 
#                names_to = "gran", 
#                values_to = "categories")

moy <- characterise_plot_data %>%
  ggplot(aes(x=month_year, y = general_supply_kwh)) +
  geom_boxplot() +
  facet_wrap(~group, ncol = 1, scales = "free_y")

hod <- characterise_plot_data %>%
  ggplot(aes(x=hour_day, y = general_supply_kwh)) +
  geom_boxplot() +
  facet_wrap(~group, ncol = 1, scales = "free_y")

wndwday <- characterise_plot_data %>%
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  geom_boxplot() +
  facet_wrap(~group, ncol = 1, scales = "free_y")

ggpairs_nocolor <- ggscatmat(cust_by_wpd_join_groups, columns=2:4) +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

cust_by_wpd_join_groups$group = as.factor(cust_by_wpd_join_groups$group)
```

```{r parcoord}
parcoord <- GGally::ggparcoord(cust_by_wpd_join_groups ,
                   columns = 2:4,
                   groupColumn = "group",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.2,
                   order = "anyClass",
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  #scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")
parcoord

# PCA based clustering

pc <- prcomp(cust_by_wpd_join_groups[c(-1,-5, -6)], center = F, scale = F)
summary(pc)
# First two principal componentsafc
comp <- data.frame(pc$x[,1:2])

data_pcp_all <- comp%>% as_tibble() %>% 
  bind_cols(cust_by_wpd_join_groups)

data_pcp_all %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = group)) +
  scale_color_brewer(palette = "Dark2") + theme_bw() + theme(legend.position = "bottom")
# 
# 
# O3y <- O3prep(data, method=c("HDo", "PCS", "BAC", "adjOut", "DDC", "MCD"))
# O3y1 <- O3plotM(O3y)
# cx <- data.frame(outlier_method=names(O3y1$nOut), number_of_outliers=O3y1$nOut)
# knitr::kable(cx, row.names=FALSE)
# 
# grid.arrange(O3y1$gO3, O3y1$gpcp, ncol=1)
```

```{r}
quantile_prob_val = c(0.25, 0.5, 0.75)

sm_hod_dow <- elec %>% 
  left_join(groups)

  sm_hod_dow %>% 
  filter(group==1) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 1")

  sm_hod_dow %>% 
  filter(group==2) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 2")

  sm_hod_dow %>% 
  filter(group==3) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 3")

  sm_hod_dow %>% 
  filter(group==4) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 4")

  sm_hod_dow %>% 
  filter(group==5) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 5")

  sm_hod_dow %>% 
  filter(group==6) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 6")

  sm_hod_dow %>% 
  filter(group==7) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 7")

  sm_hod_dow %>% 
  filter(group==8) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 8")

  sm_hod_dow %>% 
  filter(group==9) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 9")

  sm_hod_dow %>% 
  filter(group==10) %>% 
  gravitas::prob_plot(
     "month_year",
     "hour_day",
  response = "general_supply_kwh",
  plot_type = "quantile", nrow=1) +
  #scale_y_log10() +
    ggtitle("Group 10")
```