### 2021-09-22 supplementary/distance-computation/simulation-data/D2_I5_C1.Rmd

---
title: "2 designs and 5 iterations and categorty changes across 1 granularity (g3 and g1 categories const) "
author: "Sayani Gupta"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false
    cache: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,echo=FALSE, message=FALSE, warning = FALSE)
library(patchwork)
library(distributional)
library(ggplot2)
library(tidyverse)
library(knitr)
library(readr)
library(GGally)
library(tsibble)
library(tidyverse)
library(gracsr)
library(gravitas)

```

# _Simulation design_

```{r change-var, echo=TRUE}
t = 300
n <- seq(0, t, 1)
g1 <- n %%2
g2 <- n %%3
g3 <- n %%5
mu1= c(0, 0) # mean of cat of g1
mu2 = c(7, 5, 1)# mean of cat of g2
mu3 = c(0, 0, 0, 0, 0) # mean of cat of g3
```

Three circular granularities g1, g2 and g3 are considered with levels 2, 3 and 5 respectively. Many time series with `r t` observations are created using the five designs below, each of which is iterated five times. We anticipate to have 3 clusters, each with five time series conforming to the same design, once we execute the clustering.

# _Distance computation_

1) wpd for three granularities computed and then distance between designs are computed by computing the euclidean distances between them - gran as variables, 
designs across rows and cell values as wpd.

2) Robust scaling done for each customer and then JSD is computed between same categories for a granularity.

# _Granularities design_
```{r}
design = c("design-1", "design-2")
g1 = c("no", "no")
g2 = c("no", "yes")
g3 = c("no", "no")
table <- tibble(design, g1, g2, g3) %>% kable()
table

```

# _Simulate data_

```{r generate-design}
generate_design <- function(t, mu1, mu2, mu3){
t <- seq(0, t, 1)
g1 <- t %%2
g2 <- t %%3
g3 <- t %%5

# null design
g1_dnull <- rep( rep(0, each = length(unique(g1))), length.out= length(t))
g2_dnull <- rep( rep(0, each = length(unique(g2))), length.out= length(t))
g3_dnull <- rep( rep(0, each = length(unique(g3))), length.out= length(t))

# mean changing across categories in varying ways

g1_dvary <- rep(mu1, length.out= length(t))
g2_dvary <- rep(mu2, length.out= length(t))
g3_dvary <- rep(mu3, length.out= length(t))

design1 = distributional::dist_normal(g1_dnull + g2_dnull + g3_dnull)
design2 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
#design3 = distributional::dist_normal(g1_dnull + g2_dvary + g3_dnull)
#design4 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dnull)
#design5 = distributional::dist_normal(g1_dvary + g2_dvary + g3_dvary)

data_bind <- tibble::tibble(
index = t,
g1 = g1,
g2 = g2,
g3 = g3,
design1 = distributional::generate(design1, times = 1) %>% unlist(),
design2 = distributional::generate(design2, times = 1) %>% unlist(),
#design3 = distributional::generate(design3, times = 1) %>% unlist(),
#design4 = distributional::generate(design4, times = 1) %>% unlist(),
#design5 = distributional::generate(design5, times = 1) %>% unlist()
) %>% 
  pivot_longer(-c(1, 2, 3, 4), names_to = "design", values_to = "sim_data")

data_bind
}

data_bind <- generate_design(t, mu1, mu2, mu3)

data_bind

```

# _Plot raw data_

```{r plot-linear , out.width="100%"}

# plot_linear_data <- function(data){
# ggplot(data,
#              aes(x = index, y = sim_data)) + 
#   geom_line() +
#   xlab("index")+
#   theme_bw() 
# }

ggplot(data_bind,
             aes(x = index, y = sim_data)) + 
  geom_line() +
  xlab("index")+
  theme_bw() +
  facet_wrap(~design, scales = "free_y",ncol =1)

```

# _Plot distribution across granularities_

```{r plot-gran, out.width="100%"}

p2 <- ggplot(data_bind,
             aes(x = as.factor(g1), y = sim_data)) + 
  geom_boxplot(alpha =0.5) + xlab("g1") + 
  theme_bw() +
  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue")

p3 <- ggplot(data_bind, aes(x = as.factor(g2), y = sim_data)) + geom_boxplot(alpha =0.5) + xlab("g2") + theme_bw() +
  facet_wrap(~design, scales = "free_y",ncol = 1)+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue")

p4 <- ggplot(data_bind, aes(x = as.factor(g3), y = sim_data)) + geom_boxplot(alpha =0.5) +
  xlab("g3") + theme_bw()+
  facet_wrap(~design, scales = "free_y", ncol = 1)+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue")

(p2 + p3 + p4) * theme(
    strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))

# 
# 
# plot_cyclic_data( data_bind %>% filter(design=="design2"))
# plot_data( data_bind %>% filter(design=="design3"))
# plot_data( data_bind %>% filter(design=="design4"))
# plot_data( data_bind %>% filter(design=="design5"))
```

```{r,eval=FALSE}

data_bind_all <- data_bind %>% pivot_longer(-c(index, design, sim_data), names_to = "gran", values_to = "categories")

ggplot(data_bind_all, aes(x = as.character(categories), y = sim_data)) + geom_boxplot() + facet_grid(design~gran, scales = "free_y")

```

# _Iterate designs_

```{r}
seed_len = seq(12345, 12349, 1)
bind_data_iter <- map(seq_len(length(seed_len)), function(x){
set.seed(seed_len[x])
data_bind <- generate_design(t, mu1, mu2, mu3)
}) %>% bind_rows(.id = "seed_id") %>% 
  mutate(customer_id = paste(design,seed_id, sep ="-"))

bind_data_iter_tsibble <- bind_data_iter %>% 
  tsibble::as_tsibble(index = index, key = customer_id)

bind_data_iter_tsibble
```

# _Cluster designs: distance using wpd_

```{r wpd}
harmony_tbl <- tibble(facet_variable = NA, x_variable = c("g1", "g2", "g3"),
                      x_levels = c(2, 3, 5), facet_levels = NA)

dist_mat <- bind_data_iter_tsibble %>%
  #scale_gran(method = "robust", response = "sim_data") %>%
  dist_wpd(harmony_tbl, response = "sim_data", nperm=100)

groups = dist_mat%>% clust_gran(kopt = 2 )
pred_group = paste("design",groups$group,sep = "") %>% as.factor()
actual_group = as.factor(bind_data_iter_tsibble %>%as_tibble %>% select(customer_id, design) %>% distinct() %>% pull(design))

xtab <- caret::confusionMatrix(pred_group, actual_group)
xtab$table
```

# _Cluster designs: distance using js distance between categories_

```{r js}
dist_mat_g1 <- bind_data_iter_tsibble %>%
  scale_gran(method = "robust", response = "sim_data") %>%
  dist_gran(gran1 = "g1", response= "sim_data")

dist_mat_g2 <- bind_data_iter_tsibble %>%
  scale_gran(method = "robust", response = "sim_data") %>%
  dist_gran(gran1 = "g2", response= "sim_data")

dist_mat_g3 <- bind_data_iter_tsibble %>%
  scale_gran(method = "robust", response = "sim_data") %>%
  dist_gran(gran1 = "g3", response= "sim_data")

dist_mat <- dist_mat_g1 + dist_mat_g2 + dist_mat_g3

groups = dist_mat %>% 
  clust_gran(kopt = 2 )

pred_group = paste("design",groups$group,sep = "") %>% as.factor()
actual_group = as.factor(bind_data_iter_tsibble %>%as_tibble %>% select(customer_id, design) %>% distinct() %>% pull(design))

xtab <- caret::confusionMatrix(pred_group, actual_group)
xtab$table
```

# _Cluster designs: distance using js distance between categories (nqt)_

```{r js-nqt}
dist_mat_g1 <- bind_data_iter_tsibble %>%
  scale_gran(method = "nqt", response = "sim_data") %>%
  dist_gran(gran1 = "g1", response= "sim_data")

dist_mat_g2 <- bind_data_iter_tsibble %>%
  scale_gran(method = "nqt", response = "sim_data") %>%
  dist_gran(gran1 = "g2", response= "sim_data")

dist_mat_g3 <- bind_data_iter_tsibble %>%
  scale_gran(method = "nqt", response = "sim_data") %>%
  dist_gran(gran1 = "g3", response= "sim_data")

dist_mat <- dist_mat_g1 + dist_mat_g2 + dist_mat_g3

groups = dist_mat %>% 
  clust_gran(kopt = 2 )

pred_group = paste("design",groups$group,sep = "") %>% as.factor()
actual_group = as.factor(bind_data_iter_tsibble %>%as_tibble %>% select(customer_id, design) %>% distinct() %>% pull(design))

xtab <- caret::confusionMatrix(pred_group, actual_group)
xtab$table
```