### 2021-09-22 supplementary/effect_transformation.Rmd

---
title: "check_effect_transformation"
author: "Sayani Gupta"
date: "01/09/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(gravitas)
library(tidyverse)
library(gracsr)
library(tsibble)

# Feed in data and other inputs
sm <- smart_meter10 %>%
filter(customer_id %in% c("10006704", "10017936","10006414", "10018250"))
gran1 = "hour_day"
gran2 = NULL
response = "general_supply_kwh"

# Scale the data

v2 <- suppressWarnings(robust_scale_data(sm, "hour_day")) %>% 
    dplyr::rename( "kwh_robust" = "scaled_response") %>% 
    dplyr::mutate(kwh_nqt = stats::qqnorm(general_supply_kwh, plot.it=FALSE)$x)

C

quantile_q2 <-  function(x){
  y = quantile(x, probs = c(0.5))
  #c(y[1], y[2]) %>% as_tibble() %>% bind_cols(names(y)) %>% set_names(c("quant_value", "quantile"))
}

quantile_q1 <-  function(x){
  y = quantile(x, probs = c(0.25))
  #c(y[1], y[2]) %>% as_tibble() %>% bind_cols(names(y)) %>% set_names(c("quant_value", "quantile"))
}

quantile_q3 <-  function(x){
  y = quantile(x, probs = c(0.75))
  #c(y[1], y[2]) %>% as_tibble() %>% bind_cols(names(y)) %>% set_names(c("quant_value", "quantile"))
}

v2 %>%
dplyr::filter(customer_id %in% c("10018250")) %>%
  pivot_longer(c("general_supply_kwh", "kwh_robust", "kwh_nqt"),
               names_to = "kwh_type",
               values_to = "kwh_value") %>% ggplot(fill = "#999999") +
  ggridges::geom_density_ridges(aes(x = kwh_value, y = as.factor(category)), alpha = 0.7) +
  facet_grid(kwh_type~customer_id, scales = "free") +
  coord_flip() +
  theme(legend.position = "bottom") +
  theme_bw()

v2 %>%
dplyr::filter(customer_id %in% c("10018250")) %>%
  pivot_longer(c("general_supply_kwh", "kwh_robust", "kwh_nqt"),
               names_to = "kwh_type",
               values_to = "kwh_value") %>%
ggplot(aes(x = kwh_value, y = as.factor(category)), fill = "#999999") +
  #ggridges::geom_density_ridges(alpha = 0.7) +
  facet_grid(kwh_type~customer_id, scales = "free") +
  coord_flip() +
  stat_summary(
    fun = quantile_q2,
    geom = 'line',
    aes(group = 1), size = 1, color = "blue") +
  theme(legend.position = "bottom") +
  stat_summary(
    fun = quantile_q1,
    geom = 'line',
    aes(group = 1), size = 1, color = "darkgreen") +
  theme(legend.position = "bottom") +
  stat_summary(
    fun = quantile_q3,
    geom = 'line',
    aes(group = 1), size = 1, color = "#D55E00") +
  theme(legend.position = "bottom") +
  theme_bw()

# library(tidyverse)
#
# # for loop
#
# for (x in 1:2){
#   for(y in 1:2){
#     for(z in 1:5){
#       dist_data[x, y] = x*y + y*z
#     }
#   }
#
# }
#
#
#
# tab <- expand.grid(x = 1:2, y = 1:2, z =1:5)
# tab
#
# # Using map
#
#
#
# # using pmap
#
# dist_data <- purrr::pmap(tab,
#                                   function(x, y, z){
#                                     value3 =
#                                       x*y + y*z
# }) %>%
#   unlist () %>%
#   as_tibble() %>%
#   bind_cols(tab)

```