### 2021-10-07 supplementary/distance-computation/real-data/04_handpick_from356-mds_nqt.Rmd
intro applcation ssection

---
title: "Selecting similar behaving group of customers to check nqt clustering methodology"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE, message=FALSE, warning=FALSE)
library(readr)
library(tidyverse)
library(gravitas)
library(tsibble)
library(parallel)
library(magrittr)
library(GGally)
library(knitr)
library(patchwork)
library(caret)
library(plotly)
library(gracsr)
library(here)
```

# Data

Consider customers for which moy, hod and wknd_wday are all harmonies from an initial sample of 600 customers which contained no missing values. Also consider customers for which the quantiles are of same value (to avoid problems in JS computation).

```{r}
elec_356_raw <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))

elec_356_raw
```

# Classical MDS from distance metrics

```{r read-data}
# script to assimilate results from 02_algo1_moy_356cust_robust.R, 02_algo1_hod_356cust_robust.R and 02_algo1_wkndwday_356cust_robust.R
# write code after running the results for each of the above mentioned code

library(tidyverse)
library(gravitas)
library(tsibble)
library(ggpubr)

quantile_prob_graph <- c(0.25, 0.5, 0.75)

wkndwday <- read_rds("../../../data/dist_gran_wkndwday_356cust_nqt.rds") %>% broom::tidy()

moy <- read_rds("../../../data/dist_gran_moy_356cust_nqt.rds") %>% broom::tidy()

hod <- read_rds("../../../data/dist_gran_hod_356cust_nqt.rds") %>% broom::tidy()

distance <- wkndwday %>% 
  left_join(moy, by = c("item1", "item2")) %>% 
  left_join(hod, by = c("item1", "item2")) %>% 
  rename("wkndwday" ="distance.x",
         "moy" = "distance.y",
         "hod" = "distance") %>%
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

create_mds_gran <- function(total_distance){

total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
                               names_from = item2,
                               values_from = total)

rownames(total_distance_wide) <- total_distance_wide$item1

#  changing to a lower traingular matrix as dist only takes lower traingular matrix (so in way tranposing it but tidyverse style, also creating a diagonal 0 with DM otherwise as.dist changing value)
mds_data <- total_distance_wide %>% 
  mutate_all(~replace(., is.na(.), 0)) %>%
   tibble::rownames_to_column() %>%  
   select(-item1) %>% 
   pivot_longer(-rowname) %>% 
   pivot_wider(names_from=rowname, values_from=value)

df <- mds_data[-1] %>% as.matrix()
DM <- matrix(0, ncol(mds_data), ncol(mds_data))
DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)]

f = as.dist(DM)

first_lot <- mds_data %>% names()
id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])

mds  = cmdscale(f)
rownames(mds) = id
colnames(mds) = c("mds1", "mds2")

mds_hod <- bind_cols(mds, customer_id = id)

mds_hod
 }

# for hod MDS scale

total_distance_hod <- distance %>% 
  mutate(total = hod)

mds_hod <- create_mds_gran(total_distance_hod)

colnames(mds_hod) <- c("mds1.hod", "mds2.hod", "customer_id")

# for moy MDS scale

total_distance_moy <- distance %>% 
  mutate(total = moy)

mds_moy <- create_mds_gran(total_distance_moy)

colnames(mds_moy) <- c("mds1.moy", "mds2.moy", "customer_id")

# for wkndwday MDS scale

total_distance_wkndwday <- distance %>% 
  mutate(total = wkndwday)

mds_wkndwday <- create_mds_gran(total_distance_wkndwday)

colnames(mds_wkndwday) <- c("mds1.wkndwday", "mds2.wkndwday", "customer_id")

# for total MDS scale

total_distance_total <- distance %>% 
  mutate(total = wkndwday/2 + moy/12 + hod/24)

mds_total <- create_mds_gran(total_distance_total)

colnames(mds_total) <- c("mds1.total", "mds2.total", "customer_id")

```

```{r wpd-mds}

## wpd

elec_356_raw <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))

elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))

elec_pick <- elec_600_wpd %>% 
  filter(customer_id %in% elec_356_raw$customer_id)

elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)

mds <- elec_pick_wide[-1] %>% dist() %>% cmdscale() %>% as_tibble()

colnames(mds) <- c("mds1.wpd", "mds2.wpd")

mds_wpd <- bind_cols(mds, customer_id = elec_pick_wide$customer_id)%>% mutate(customer_id = as.character(customer_id))
```

```{r mds-all}
mds_all <- mds_hod %>% 
  left_join(mds_moy, by = "customer_id") %>% 
  left_join(mds_wkndwday, by = "customer_id") %>% 
  left_join(mds_total, by = "customer_id") %>% 
  left_join(mds_wpd, by = "customer_id") %>% 
  select(customer_id, everything())

```

# Visualize similarity between samples for different distance metrics hod+moy+wkndwday, hod, moy, wkndwday

```{r mds-all-2}

mds_tx <- highlight_key(mds_all, ~customer_id)

# initiate a plotly object
base <- plot_ly(mds_tx, color = I("black")) %>% 
  group_by(customer_id)

mds_hod <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id)

mds_moy <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.moy, y = ~ mds2.moy, text = ~customer_id)

mds_wkndwday <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.wkndwday, y = ~ mds2.wkndwday, text = ~customer_id)

mds_total <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.total, y = ~ mds2.total, text = ~customer_id)

mds_wpd <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.wpd, y = ~ mds2.wpd, text = ~customer_id)

subplot(mds_total, mds_hod, mds_moy ,mds_wkndwday, titleX = TRUE) %>%
  layout(showlegend = FALSE) %>%
  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect") 
```

# Visualize similarity between js and wpd methodologies:

```{r wpd-total}
subplot(mds_wpd, mds_total, nrows = 1, titleX = TRUE) %>%
  layout(showlegend = FALSE) %>%
  highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = 0.2, off = "plotly_deselect")
```

## --- Total

```{r interactive-characterization}

quantile_hod <- quantile_gran(elec_356_raw, gran1 = "hour_day", quantile_prob_val = quantile_prob_graph)

quantile_moy <- quantile_gran(elec_356_raw, gran1 = "month_year", quantile_prob_val = quantile_prob_graph)

quantile_wkndwday <- quantile_gran(elec_356_raw, gran1 = "wknd_wday", quantile_prob_val = quantile_prob_graph)

quantile_data <- bind_rows(quantile_hod, quantile_moy, quantile_wkndwday)

elec_pick_wide <- elec_pick %>% pivot_wider(c(3:5), names_from = "x_variable", values_from = wpd)

data_heatmap <- quantile_data %>% 
 left_join(elec_pick_wide) %>% 
  mutate(customer_id = as.character(customer_id))

# %>% 
#   mutate(categories_serial_id = as.factor(categories_serial_id)) %>% 
#   mutate(customer_id = as.character(customer_id))

data_heatmap_new <- data_heatmap %>%
left_join(mds_all, by = "customer_id") %>% 
  pivot_wider( names_from = quantiles, 
               values_from = quantiles_values)

tx <- highlight_key(data_heatmap_new,
                    ~customer_id)

 # initiate a plotly object
 base <- plot_ly(tx, color = I("black")) %>%
   group_by(customer_id)
#
 quantile_plot <- base %>% 
   group_by(customer_id, gran) %>%
   add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
   add_ribbons(x = ~category,
               ymin = ~`25%`,
               ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
   layout(showlegend = FALSE)

mds_total <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.total, y = ~ mds2.total, text = ~customer_id)

mds_wpd <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.wpd, y = ~ mds2.wpd, text = ~customer_id)

subplot(mds_total, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE)
```

# hod

```{r}
data_heatmap_hod <- data_heatmap_new %>% 
  filter(gran == "hour_day") %>% 
  mutate(category = as.integer(category))

data_heatmap_hod$category <- factor(data_heatmap_hod$category, levels = 0:23)

tx <- highlight_key(data_heatmap_hod, ~customer_id)

 # initiate a plotly object
base <- plot_ly(tx, color = I("black")) %>%
   group_by(customer_id)

 quantile_plot <- base %>%
   group_by(customer_id, gran) %>%
   add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
   add_ribbons(x = ~category,
             ymin = ~`25%`,
             ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
   layout(showlegend = FALSE)

 mds_hod <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id)

subplot(mds_hod, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE)
#

```

# moy

```{r}

data_heatmap_moy <- data_heatmap_new %>% 
  filter(gran == "month_year")

data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

#%>%  mutate(category = as.integer(category))

tx <- highlight_key(data_heatmap_moy, ~customer_id)

 # initiate a plotly object
base <- plot_ly(tx, color = I("black")) %>%
   group_by(customer_id)

 quantile_plot <- base %>%
   group_by(customer_id, gran) %>%
   add_lines(x = ~ category, y = ~`50%`, color = ("black")) %>%
   add_ribbons(x = ~category,
             ymin = ~`25%`,
             ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>%
   layout(showlegend = FALSE)

 mds_moy <- base %>%
  group_by(customer_id) %>%
  add_trace(x = ~mds1.moy, y = ~ mds2.moy, text = ~customer_id)

subplot(mds_moy, quantile_plot, widths = c(.6, .4), titleX = TRUE) %>% layout(showlegend = FALSE) %>% highlight(on = "plotly_selected", dynamic = TRUE, selectize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE)
```

## run upto this

```{r, eval = FALSE}

distance <- wkndwday %>% 
  left_join(moy, by = c("item1", "item2")) %>% 
  left_join(hod, by = c("item1", "item2")) %>% 
  rename("wkndwday" ="distance.x",
         "moy" = "distance.y",
         "hod" = "distance") %>% mutate(item1 = as.integer(as.character(item1)))

#10850478 8486921
data_pick_one <- total_distance %>% filter(item1 %in% 
                            c(8454221)) %>% group_by(item1) %>% arrange(total) %>% head(5) %>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

#11013154  8181071 11048034
data_pick_two <- total_distance %>% filter(item1 %in% 
                            c(8181071)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

data_pick_three <- total_distance %>% filter(item1 %in% 
                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

data_pick_three <- total_distance %>% filter(item1 %in%
                            c(11013154)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

data_pick_four <- total_distance %>% filter(item1 %in%
                            c(8653709)) %>% group_by(item1) %>% arrange(total) %>% head(5)%>%
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))
```

```{r,  eval = FALSE}

data_pick_cust <- bind_rows(
unique(c(data_pick_one$item1,data_pick_one$item2)) %>% as_tibble(),    unique(c(data_pick_two$item1,data_pick_two$item2)) %>% as_tibble(), c(11013154, 8495194, 8627007,10109182,10677705,8952846) %>% as_tibble(),unique(c(data_pick_four$item1,data_pick_four$item2)) %>% as_tibble(),
.id = "design")

data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
  filter(customer_id %in% data_pick_cust$value) %>% 
  gracsr::scale_gran( method = "robust", response = "general_supply_kwh")

sm_hod <- gravitas::create_gran(data_pick, "hour_day")

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)  %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value"))

ncol_sm <- seq_len(ncol(sm_hod_list[-c(1, ncol(sm_hod_list))]))
nrow_sm <- seq_len(nrow(sm_hod_list))

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list%>% 
     dplyr::filter(customer_id == customer_id[x]) %>% 
     select(-c(1, ncol(sm_hod_list))) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_graph, na.rm = TRUE)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "customer_serial_id")

ref_cat <- names(sm_hod_list)[-c(1, ncol(sm_hod_list))] %>% as_tibble() %>% set_names("category") %>%  
  mutate(categories_serial_id = row_number())

ref_cust <- sm_hod_list$customer_id %>% as_tibble()%>% set_names("customer_id")  %>% 
    mutate(customer_serial_id = row_number())

sm_quantiles_ref <- sm_hod_quantiles_cat %>% 
  mutate(customer_serial_id = as.integer(customer_serial_id),
         categories_serial_id = as.integer(categories_serial_id)) %>%
  left_join(ref_cat, by = "categories_serial_id") %>% 
  left_join(ref_cust, by = "customer_serial_id") %>% 
  select(customer_serial_id, categories_serial_id, category, customer_id, everything())

data_heatmap <- sm_quantiles_ref %>% 
 left_join(data_pick %>% distinct(customer_id)) %>% 
  left_join(data_pick_cust, by = c("customer_id" = "value")) %>% 
  mutate(categories_serial_id = as.factor(categories_serial_id))

data_heatmap %>% 
  ggplot(aes(x = categories_serial_id)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id)) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + ylab("demand (in Kwh)") + theme_bw() + theme(panel.spacing =unit(0, "lines")) + theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
  scale_x_discrete(breaks = seq(1, 24, 3))

```
<!-- # # client side linking -->
<!-- # ```{r} -->
<!-- # mds <- bind_cols(mds, customer_id = total_distance_wide$item1) -->
<!-- #  -->
<!-- # data_heatmap_new <- data_heatmap %>%  -->
<!-- #   left_join(mds, by = "customer_id") -->

<!-- tx <- highlight_key(data_heatmap_new, ~customer_id) -->

<!--  # initiate a plotly object -->
<!--  base <- plot_ly(tx, color = I("black")) %>% -->
<!--    group_by(customer_id) -->
<!-- # -->
<!--  quantile_plot <- base %>% -->
<!--    group_by(customer_id) %>% -->
<!--    add_lines(x = ~ categories_serial_id, y = ~`50%`, color = ("black")) %>% -->
<!--    add_ribbons(x = ~categories_serial_id, -->
<!--                ymin = ~`25%`, -->
<!--                ymax = ~`75%`, color = I("lightblue"), alpha = (0.5))%>% -->
<!--    layout(showlegend = FALSE) -->

<!--  # fig <- plot_ly(base,color = I("blue")) %>% roup_by(customer_id) -->

<!--  fig_trace <- base %>% -->
<!--    group_by(customer_id) %>% -->
<!--    add_trace(x = ~Dim.1, y = ~ Dim.2, text = ~customer_id) -->

<!--  subplot(fig_trace, quantile_plot, widths = c(.6, .4), titleX  TRUE) %>% -->
<!--    layout(showlegend = FALSE) %>% -->
<!--    highlight(on = "plotly_selected", dynamic = TRUE, electize = TRUE, selected = attrs_selected(0.5), opacityDim = .2, off = "plotly_deselect", persistent = TRUE, defaultValues  c(8271504, 9316697)) -->

<!--  %>% -->
<!--    highlight(color = toRGB("black")) -->

<!--              , color = list( -->
<!--    toRGB("red", "black") -->
<!--    ) -->
<!--  ) -->

<!-- # mds_hod <- base %>% -->
<!--   # group_by(customer_id) %>% -->
<!--   # add_trace(x = ~mds1.hod, y = ~ mds2.hod, text = ~customer_id) -->

<!-- # hod_wide <- hod %>% -->
<!-- #   pivot_wider(names_from = item2, values_from = distance)%>% -->
<!-- #   mutate(item1 = as.integer(as.character(item1))) -->
<!-- #  -->
<!-- # hod_wide$`8143599` <- NA -->
<!-- #  -->
<!-- # hod_wide <- bind_rows(hod_wide, c(item1=11590442)) -->
<!-- #  -->
<!-- # mds_data <- hod_wide[-1]  %>% mutate_all(~replace(., is.na(.), 0)) %>% as.dist() %>% clust_gran() -->
<!-- #   -->