### 2021-10-18 check_work.Rmd
adding new simulation files

---
title:  Selecting customers

abstract: |
 Clustering elements based on behavior across time granularities 
bibliography: [bibliography.bib]
preamble: >
  \setlength {\marginparwidth }{2cm}
  \usepackage{mathtools,amssymb,booktabs,longtable,todonotes,amsthm}
  \def\mod{~\text{mod}~}
output:
  bookdown::pdf_book:
    base_format: rticles::asa_article
    fig_height: 4
    fig_width: 6
    fig_caption: yes
    dev: "pdf" 
    keep_tex: yes
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsr)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
library(patchwork)
library(GGally)
library(distributional)
library(viridis)
```

```{r mytheme}
theme_characterisation <- function() {

  theme_bw() + # seeting theme
    theme(strip.text = element_text(size = 9,
                                    margin = margin(b = 0, t = 0))) + # narrow facet space
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + # no axis ticks 
  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
    theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
    theme(legend.position = "bottom")+
  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
  theme(axis.text.x = element_text(size=5))
}
```

```{r anchor}
anchor1 <- 11156027
anchor2 <- 10420683
anchor3 <- 8629763
anchor4 <- 11013154
kopt <- 4 # number of clusters
```

_Where they lie on the mds plot of total, hod, moy, wkndwday and wpd_

```{r data-read}

quantile_prob_graph <- c(0.25, 0.5, 0.75)

wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_robust.rds"))

moy <- read_rds(here("data/dist_gran_moy_356cust.rds"))

hod <- read_rds(here("data/dist_gran_hod_356cust_robust.rds"))

distance <- wkndwday %>% 
  left_join(moy, by = c("item1", "item2")) %>% 
  left_join(hod, by = c("item1", "item2")) %>% 
  rename("wkndwday" ="distance.x",
         "moy" = "distance.y",
         "hod" = "distance") %>%
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2))) 
```

```{r data-pick}
total_distance <- distance %>% 
  #mutate(total = hod)
  mutate(total = wkndwday/2 + moy/12 + hod/24)

# 8454221 10420689 1001 8454221 10018298 10420689 9393696

# k-nearest neighbor pick function
nn_pick_function <- function(anchor, k = 5){
  data <- total_distance %>% filter(item1 %in% 
                            c(anchor)|item2 %in% 
                            c(anchor) ) %>% 
  #group_by(item1) %>%
  arrange(total) %>% 
  head(k) %>% 
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

  unique(c(data$item1,data$item2)) %>% as_tibble()
}

data_pick_one <- nn_pick_function(anchor1)
data_pick_two <- nn_pick_function(anchor2)
data_pick_three <- nn_pick_function(anchor3)
data_pick_four <- nn_pick_function(anchor4)

# c(11013154, 9345642, 8328008, 8636035,8627007,8454235)

data_pick_cust <- bind_rows(
data_pick_one, data_pick_two, data_pick_three, data_pick_four,
.id = "design")

data_pick_cust$design <- paste("design", data_pick_cust$design, sep = "-")

data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
  filter(customer_id %in% data_pick_cust$value) %>% 
  gracsr::scale_gran(method = "robust", response = "general_supply_kwh")
```

```{r mds-plots}

data_pick_char <- data_pick_cust %>% 
  rename("customer_id" = "value") %>% 
  mutate(customer_id = as.character(customer_id))

mds_all <- read_rds(here("data/mds_all.rds"))  %>% 
  left_join(data_pick_char)

highlight_df <- mds_all %>% filter(customer_id %in% data_pick_char$customer_id)

total_plot <- ggplot(mds_all,
aes(x = mds1.total, y = mds2.total)) +
  geom_point(alpha = 0.3) +
  geom_point(data = highlight_df, 
             aes(x = mds1.total, y = mds2.total, color = as.factor(design)), size = 2)

wpd_plot <- ggplot(mds_all,
aes(x = mds1.wpd, y = mds2.wpd)) +
  geom_point(alpha = 0.3) +
  geom_point(data = highlight_df, 
             aes(x = mds1.wpd, y = mds2.wpd, color = as.factor(design)), size = 2)

hod_plot <- ggplot(mds_all,
aes(x = mds1.hod, y = mds2.hod)) +
  geom_point(alpha = 0.3) +
  geom_point(data = highlight_df, 
             aes(x = mds1.hod, y = mds2.hod, color = as.factor(design)), size = 2)

moy_plot <- ggplot(mds_all,
aes(x = mds1.moy, y = mds2.moy)) +
  geom_point(alpha = 0.3) +
  geom_point(data = highlight_df, 
             aes(x = mds1.moy, y = mds2.moy, color = as.factor(design)), size = 2)

wkndwday_plot <- ggplot(mds_all,
aes(x = mds1.wkndwday, y = mds2.wkndwday)) +
  geom_point(alpha = 0.3) +
  geom_point(data = highlight_df, 
             aes(x =  mds1.wkndwday, 
                 y =  mds2.wkndwday, 
                 color = as.factor(design)),
             size = 2)

p <- (total_plot + wpd_plot)/(hod_plot + moy_plot + wkndwday_plot)

p* +
  plot_layout(guides = "collect")

```

```{r clustering}
filtered_distance <- distance %>% filter(item1 %in% data_pick_cust$value) %>% 
  filter(item2 %in% data_pick_cust$value)

total_distance <- filtered_distance %>% 
mutate(total = wkndwday/2 + moy/12 + hod/24) 
#mutate(total = hod)

total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
                               names_from = item2,
                               values_from = total)
#total_distance_wide$`11013154` <- NA

rownames(total_distance_wide) <- total_distance_wide$item1

mds_data <- total_distance_wide %>% 
  mutate_all(~replace(., is.na(.), 0)) %>%
   tibble::rownames_to_column() %>%  
   select(-item1) %>% 
   pivot_longer(-rowname) %>% 
   pivot_wider(names_from=rowname, values_from=value)

rownames(mds_data) <- total_distance_wide$item1

# group <- mds_data[-1] %>% clust_gran(kopt=4)
#

df <- mds_data[-1] %>% as.matrix()
DM <- matrix(0, ncol(mds_data), ncol(mds_data))
DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)]

f = as.dist(DM)

#f %>% hclust %>% cutree(k=4)

first_lot <- mds_data %>% names()
# because distance matrix is such  that 1:23 appears in columns and 2:24 appears in rows such that 1st customer is missing from row and last customer is missing from variables
id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])

group <- f %>% hclust (method = "ward.D") %>% cutree(k=kopt)

cluster_result <- bind_cols(id = id, group = group)

mds  = cmdscale(f)
rownames(mds) = id
colnames(mds) = c("mds1", "mds2")
```

```{r heatmap-hod}

data_heatmap_hod <- quantile_gran(data_pick,
                                  "hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles,
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"="value"))

#data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
data_heatmap_hod$category <- factor(data_heatmap_hod$category, levels = 0:23)

```

```{r data-heatmap-hod-new}

legend_title <- "group"

data_heatmap_hod %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id,
                  fill = as.factor(group)),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id, 
                color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) + theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_x_discrete(breaks = seq(0, 23, 3))
# + scale_y_continuous(breaks = NULL) 
```

```{r data-heatmap-moy}

data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"="value"))

data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, 
              fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year") +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
```

```{r data-heatmap-wkndwday}

data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"="value"))

data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))

data_heatmap_wkndwday %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, 
              fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
 theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
```

```{r data-heatmap-hod-group}

data_group <- data_pick %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id"))

data_heatmap_hod_group <- quantile_gran(data_group,
                                  gran1="hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)

# data_heatmap_hod_group <- data_heatmap_hod_group %>%
#   mutate(group = case_when(group==1 ~"Group 2",
#                            group==2 ~"Group 4",
#                            group==3 ~"Group 1",
#                            group==4 ~"Group 3"))

# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))

data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")

hod_group <- data_heatmap_hod_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=group,
                  fill = as.factor(group), alpha = 0.5),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=group, 
                color = as.factor(group)), size = 1)+
  facet_wrap(~group, 
             scales = "free_y",  
             nrow = 4) + 
             theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) 
#

# 
# data_pick_cust <- data_pick_cust %>% mutate(value = as.character(value))
# 
# data_pick_cust %>% left_join(cluster_result, by = c("value" = "id")) %>% group_by(group)
```

```{r data-heatmap-moy-group}

data_group <- data_pick %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id"))

data_heatmap_moy_group <- quantile_gran(data_group,
                                  gran1="month_year",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")

moy_group <- data_heatmap_moy_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 4) +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
  ylab("demand (in Kwh)") +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
```

```{r data-heatmap-wkndwday-2}

data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"="value"))

data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))

p1 <- data_heatmap_wkndwday %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6) +
   theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))

```

```{r data-heatmap-wkndwday-group}
data_group <- data_pick %>%
  mutate(customer_id = as.character(customer_id)) %>%
  left_join(cluster_result, by = c("customer_id" = "id"))

wkndwday_data<- data_group %>% create_gran("wknd_wday") 
ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]

wkndwday_group <- wkndwday_data%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  lvplot::geom_lv(aes(fill = as.factor(group), color = as.factor(group)), k=5, alpha = 0.5) +
  #geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 4)  +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))

```

```{r combined}

combined <- hod_group + moy_group + wkndwday_group & theme(legend.position = "bottom")
combined + plot_layout(guides = "collect")

```

##  wpd-based clustering

```{r ggpairs}

elec_600_wpd <- read_rds(here::here("data/algo2-cust600-wpd-rawdata.rds"))

elec_pick <- elec_600_wpd %>% 
  filter(customer_id %in% data_pick_cust$value)

elec_pick_wide <- elec_pick %>% pivot_wider(-c(1, 2), names_from = "x_variable", values_from = wpd)

```

```{r }

# robust_scale <-  function(x){
#   (x - stats::median(x))/(stats::IQR(x))
# }

#scaled_var <- apply(elec_pick_wide[-c(1)], 2, robust_scale) %>% as_tibble()

scaled_var <- elec_pick_wide

f <- elec_pick_wide[-1] %>% dist()

group <- f%>% hclust (method = "ward.D") %>% cutree(k=kopt)

cluster_result_wpd <- bind_cols(id = elec_pick_wide$customer_id, group = group)

# mds plot

# mds  = cmdscale(f)
# rownames(mds) = id
# colnames(mds) = c("mds1", "mds2")

# ggplot(data = mds %>% as_tibble() %>% mutate(group = cluster_result_wpd$group),
#        aes(x = mds1, y = mds2, color = as.factor(group))) +
#   geom_point() +
#   geom_text(aes(label = id))

data_pcp <- scaled_var %>% 
  #bind_cols(customer_id =  elec_pick_wide$customer_id) %>%
  left_join(cluster_result_wpd , by = c("customer_id" = "id")) %>% 
  select(customer_id, group, everything()) %>% 
  mutate(group = as.factor(group))

pairsplot <- ggpairs(data_pcp, 
                     columns = 3:5,
                     aes(fill=group, color = group), alpha = 0.5) +
  scale_fill_viridis_d(direction = -1) +
  scale_color_viridis_d(direction = -1) +
  theme_light()

# will give error if you have got just one observation in a cluster

# ggplot(data = data_pcp,
#        aes(x = month_year, y = wkndwday)) +
#   geom_point() +
#   geom_text(aes(label = id))

  # 
  # scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00")) +   scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00"))

parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 3:ncol(data_pcp),
                   groupColumn = "group",
                   showPoints = FALSE, 
                   alphaLines = 0.8,
                   order = "anyClass",
                   scale = "globalminmax"
) + theme_characterisation()+
  ylab("wpd") + scale_fill_viridis_d(direction = -1) +
  scale_color_viridis_d(direction = -1) + theme_light()
```

```{r parcoord}
parcoord
```

```{r data-heatmap-hod-group-wpd}

cluster_result_wpd$id <- as.character(cluster_result_wpd$id)

data_group <- data_pick %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result_wpd, by = c("customer_id" = "id"))

data_heatmap_hod_group <- quantile_gran(data_group,
                                  gran1="hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)

# data_heatmap_hod_group <- data_heatmap_hod_group %>%
#   mutate(group = case_when(group==1 ~"Group 2",
#                            group==2 ~"Group 4",
#                            group==3 ~"Group 1",
#                            group==4 ~"Group 3"))

# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))

data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")

hod_group_wpd <- data_heatmap_hod_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=group,
                  fill = as.factor(group), alpha = 0.5),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=group, 
                color = as.factor(group)), size = 1)+
  facet_wrap(~group, 
             scales = "free_y",  
             nrow = 4) + 
              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
  ylab("demand (in Kwh)") + 
  theme_light()+ 
  theme_characterisation() +
 scale_fill_viridis_d(direction = -1) +
  scale_color_viridis_d(direction = -1)

#

# 
# data_pick_cust <- data_pick_cust %>% mutate(value = as.character(value))
# 
# data_pick_cust %>% left_join(cluster_result, by = c("value" = "id")) %>% group_by(group)
```

```{r data-heatmap-moy-group-wpd}

data_group <- data_pick %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result_wpd, by = c("customer_id" = "id"))

data_heatmap_moy_group <- quantile_gran(data_group,
                                  gran1="month_year",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")

moy_group_wpd <- data_heatmap_moy_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 4) +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
  ylab("demand (in Kwh)") + 
  theme_light()+ 
  theme_characterisation() +
 scale_fill_viridis_d(direction = -1) +
  scale_color_viridis_d(direction = -1) 
```

```{r data-heatmap-wkndwday-group-new2-wpd}
data_group <- data_pick %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result_wpd, by = c("customer_id" = "id"))

data_heatmap_wkndwday_group <- quantile_gran(data_group,
                                  gran1="wknd_wday",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_wkndwday_group$category <- factor(data_heatmap_wkndwday_group$category, levels = c("Weekday", "Weekend"))

data_heatmap_wkndwday_group$group <- paste("group", data_heatmap_wkndwday_group$group, sep = "-")

wkndwday_group_wpd <- data_heatmap_wkndwday_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 4)+
    #theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) 
    xlab("wknd-wday") + 
  ylab("demand (in Kwh)") + 
  theme_light()+ 
  theme_characterisation() +
 scale_fill_viridis_d(direction = -1) +
  scale_color_viridis_d(direction = -1)

data_group <- data_pick %>%
  mutate(customer_id = as.character(customer_id)) %>%
  left_join(cluster_result_wpd, by = c("customer_id" = "id"))

wkndwday_data<- data_group %>% create_gran("wknd_wday") 
ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]

wkndwday_group_wpd <- wkndwday_data%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  lvplot::geom_lv(aes(fill = as.factor(group), color = as.factor(group)), k=5, alpha = 0.5) +
  #geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 4)   + 
  theme_light()+ 
  theme_characterisation() +
 scale_fill_viridis_d(direction = -1) +
  scale_color_viridis_d(direction = -1) 
```

```{r combined-wpd}

combined <- hod_group_wpd + moy_group_wpd + wkndwday_group_wpd & theme(legend.position = "bottom")
combined + plot_layout(guides = "collect")

```
```{r extra}
# total
# data = mds_total %>% left_join(data_pick_char)
# 
# highlight_df <- data %>% filter(customer_id%in% data_pick_char$customer_id)
#   
# total_plot <- ggplot(data,
# aes(x = mds1.total, y = mds2.total)) +
#   geom_point(alpha = 0.3) +
#   geom_point(data = highlight_df, 
#              aes(x = mds1.total, y = mds2.total, color = as.factor(design)), size = 2)
# 
# #hod
# 
# data = mds_hod %>% left_join(data_pick_char)
# 
# highlight_df <- data %>% filter(customer_id%in% data_pick_char$customer_id)
#   
# hod_plot <- ggplot(data,
# aes(x = mds1.hod, y = mds2.hod)) +
#   geom_point(alpha = 0.3) +
#   geom_point(data = highlight_df, 
#              aes(x = mds1.hod, y = mds2.hod, color = as.factor(design)), size = 2)+theme(legend.position = "bottom")
# 
# # moy
# 
# data = mds_moy %>% left_join(data_pick_char)
# 
# highlight_df <- data %>% filter(customer_id%in% data_pick_char$customer_id)
#   
# moy_plot <- ggplot(data,
# aes(x = mds1.moy, y = mds2.moy)) +
#   geom_point(alpha = 0.3) +
#   geom_point(data = highlight_df, 
#              aes(x = mds1.moy, y = mds2.moy, color = as.factor(design)), size = 2)+theme(legend.position = "bottom")
# 
# # wkndwday
# 
# data = mds_wkndwday %>% left_join(data_pick_char)
# 
# highlight_df <- data %>% filter(customer_id %in% data_pick_char$customer_id)
#   
# wkndwday_plot <- ggplot(data,
# aes(x = mds1.wkndwday, y = mds2.wkndwday)) +
#   geom_point(alpha = 0.3) +
#   geom_point(data = highlight_df, 
#              aes(x =  mds1.wkndwday, 
#                  y =  mds2.wkndwday, 
#                  color = as.factor(design)),
#              size = 2)+ theme(legend.position = "bottom")

```