### 2021-10-26 select_cust_tour_tsne.Rmd

---
title: "Selecting customers"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsr)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
library(patchwork)
library(GGally)
library(distributional)
library(viridis)
```

```{r mytheme}
theme_characterisation <- function() {

  theme_bw() + # seeting theme
    theme(strip.text = element_text(size = 10,
                                    margin = margin(b = 0, t = 0))) + # narrow facet space
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + # no axis ticks 
  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
    theme(axis.text.x = element_text(angle=90, hjust=1, size = 10)) + # rotate the x-axis text
    theme(legend.position = "bottom")+
  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
  theme(axis.text.x = element_text(size=5))
}
```

```{r anchor}
kopt <- 3 # number of clusters
```

_How distributions look (left: tour design, right: clustering)_

```{r}
data_pick_robust <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>% 
  gracsr::scale_gran( method = "robust",
                      response = "general_supply_kwh")

# data_356cust_hod <- quantile_gran(data_pick_robust,
#               "hour_day", 
#               quantile_prob_val = seq(0.1, 0.9, 0.1))
# 
# data_356cust_moy <- quantile_gran(data_pick_robust,
#                                   "month_year", 
#                                   quantile_prob_val = seq(0.1, 0.9, 0.1))
# 
# data_356cust_wkndwday <- quantile_gran(data_pick_robust,
#                                   "wknd_wday", 
#                                   quantile_prob_val = seq(0.1, 0.9, 0.1))
#

data_356cust_hod <- read_rds("data/quantile_data_356cust_hod_robust.rds") %>% 
  filter(quantiles %in% "50%")

data_356cust_moy <- read_rds("data/quantile_data_356cust_moy_robust.rds") %>% 
  filter(quantiles %in% "50%")

data_356cust_wkndwday <- read_rds("data/quantile_data_356cust_wkndwday_robust.rds") %>% 
  filter(quantiles %in% "50%")

data_356cust_hod_wide <- data_356cust_hod %>%
  pivot_wider(names_from = c("gran", "category", "quantiles"),
              values_from = "quantiles_values")

data_356cust_moy_wide <- data_356cust_moy %>%
  pivot_wider(names_from = c("gran", "category", "quantiles"),
              values_from = "quantiles_values")

data_356cust_wkndwday_wide <- data_356cust_wkndwday %>%
  pivot_wider(names_from = c("gran", "category", "quantiles"),
              values_from = "quantiles_values")

data_356cust_wide <- left_join(data_356cust_hod_wide,
                               data_356cust_moy_wide, by="customer_id") %>% 
  left_join(data_356cust_wkndwday_wide,  by="customer_id"
                              )

save(data_356cust_wide, file="data/data_356cust_wide.rda")

# cluster_result <- cluster_result %>% mutate(customer_id = as.integer(id))

# data_356cust_wide_group <- data_356cust_wide %>% 
#   left_join(cluster_result, by = ("customer_id")) %>% 
#   mutate(group = if_else(is.na(group), 10L, group))

# data_356cust_wide_group$group <- factor(data_356cust_wide_group$group)

# without scaling

data_356cust_pc <- prcomp(data_356cust_wide[,-1],
                          center = FALSE, scale = FALSE, retx = TRUE)

#plot(data_356cust_pc, type="l", npcs=50)

# from scree-plot it looks like first 10 PCs explain most of the
# variation

library(liminal)

set.seed(2935)
data_356cust_pc10 <- as_tibble(data_356cust_pc$x[,1:6])
# %>%   mutate(group = data_356cust_wide_group$group)
#limn_tour(data_356cust_pc10, PC1:PC6)

#sort(abs(data_356cust_pc$rotation[,1]))

##---- t-SNE embeddings

library(Rtsne)
set.seed(2099)
tSNE_fit <- data_356cust_wide%>% 
  select(-customer_id) %>% 
  Rtsne( pca = FALSE,
         perplexity = 30)

data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble 
data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble
data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble
#data_pick_four <- c(10590714,8495194,8589936, 8454235) %>% as_tibble

data_pick_cust <- bind_rows(
  data_pick_one, data_pick_two, data_pick_three,
  .id = "design")

tsne_df <- data.frame(tsneX = tSNE_fit$Y[, 1], tsneY = tSNE_fit$Y[, 2], customer_id = data_356cust_wide$customer_id) %>% left_join(
  data_pick_cust, by = c("customer_id" = "value"
)) %>% mutate(design = if_else(is.na(design), "0", design))

#rownames(tsne_df) <- data_356cust_wide$customer_id

rownames(data_356cust_pc10) <- data_356cust_wide$customer_id

## ----tsne-xy--------------------------------------------------------
tsne_xy <- ggplot(tsne_df, aes(x = tsneX, y = tsneY, color = design)) +
  geom_point(aes(text = customer_id)) +
  #scale_color_manual(values = limn_pal_tableau10()) +
  scale_colour_viridis_d(direction = -1) +
  guides(color = FALSE) +
  labs(caption = "tSNE") +
  theme(aspect.ratio = 1) +
  theme_light()

## ----highlight-customer--------------------------------------------------------

library(plotly)
tsne_plotly <- tsne_xy %>% ggplotly(tooltip = "text")
```

```{r data-read}

quantile_prob_graph <- c(0.25, 0.5, 0.75)

data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble %>% set_names("customer_id")
#data_pick_four <- c(10590714,8495194,8589936, 8454235) %>% as_tibble %>% set_names("customer_id")

cust_design <- bind_rows(
data_pick_one, data_pick_two, data_pick_three,
.id = "design")

data_pick_cust <- cust_design%>% 
  mutate(customer_id = as.character(customer_id))

# data_pick_cust$design <- paste("design",
#                                data_pick_cust$design, sep = "-")
#

# original data
data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds"))%>%
  mutate(customer_id = as.character(customer_id)) %>% 
  filter(customer_id %in% data_pick_cust$customer_id)  %>% 
  gracsr::scale_gran(method = "robust", response = "general_supply_kwh")
```

```{r clustering-new}

# data_pick is already robust-scaled

hod <- data_pick %>% 
  dist_gran(gran1 = "hour_day", response = "general_supply_kwh") 
moy <- data_pick %>% 
  dist_gran(gran1 = "month_year", response = "general_supply_kwh")
wkndwday <- data_pick %>% 
  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh")

distance <- wkndwday/2 + moy/12 + hod/24

f = as.dist(distance)

f %>% hclust %>% cutree(k=3)

cluster_result <- f %>% clust_gran(kopt = kopt)

# confusion_tabl2 <- confusion_tabl1 %>% left_join(cluster_result_nobroom, by = c("customer_id" = "id"))

# broom or no-broom gives the same result
```

```{r heatmap-hod}

data_heatmap_hod <- quantile_gran(data_pick,
                                  "hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles,
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

#data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
data_heatmap_hod$category <- factor(data_heatmap_hod$category, levels = 0:23)

```

```{r data-heatmap-hod-design}

#data_pick_cust$design <- factor(data_pick_cust$design)

data_heatmap_hod_design <- data_heatmap_hod %>% 
  #mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, fill = as.factor(design)),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id, color = as.factor(design)), size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y",
             ncol = 4) + theme_characterisation()+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  scale_x_discrete(breaks = seq(0, 23, 3))
# + scale_y_continuous(breaks = NULL) 
```

```{r data-heatmap-hod-group-ind}

#data_pick_cust$design <- factor(data_pick_cust$design)

data_heatmap_hod_ind <- data_heatmap_hod %>% 
  #mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, fill = as.factor(group)),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y",
             ncol = 4) + theme_characterisation()+
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  scale_x_discrete(breaks = seq(0, 23, 3))
# + scale_y_continuous(breaks = NULL) 
# 
# 
data_heatmap_hod_design + data_heatmap_hod_ind
```

```{r data-heatmap-moy-design}

data_heatmap_moy <- quantile_gran(data_pick, "month_year", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_heatmap_moy$category <- factor(data_heatmap_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy_design <- data_heatmap_moy %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(design)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(design)), size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y", 
             ncol = 4) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_viridis_d(direction = 1)+
  scale_color_viridis_d(direction = 1)
```

```{r data-heatmap-moy-group-ind}

data_heatmap_moy_ind <- data_heatmap_moy %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(~customer_id, 
             scales = "free_y", 
             ncol = 4) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))

data_heatmap_moy_design+data_heatmap_moy_ind
```

```{r data-heatmap-wkndwday}

data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))

data_group <- data_pick %>%
  mutate(customer_id = as.character(customer_id)) %>%
  left_join(cluster_result, by = c("customer_id" = "id"))%>% 
  left_join(data_pick_cust, by = c("customer_id"))

wkndwday_data<- data_group %>% create_gran("wknd_wday") 
ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]

wkndwday_group <- wkndwday_data%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh))  +
  lvplot::geom_lv(aes(fill = design, 
                   color = design),
               alpha = 0.5)+
  coord_cartesian(ylim = ylim1*1.05)+
  # stat_summary(
  #   fun = median,
  #   geom = 'line',
  #   aes(group = 1), size = 1, color = "blue")+
  facet_wrap(~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 13) +
  theme_characterisation()+
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey"))
```

```{r data-heatmap-hod-group}

data_group <- data_pick %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id"))

data_heatmap_hod_group <- quantile_gran(data_group,
                                  gran1="hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)

# data_heatmap_hod_group <- data_heatmap_hod_group %>%
#   mutate(group = case_when(group==1 ~"Group 2",
#                            group==2 ~"Group 4",
#                            group==3 ~"Group 1",
#                            group==4 ~"Group 3"))

# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))

data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")

hod_group <- data_heatmap_hod_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=group,
                  fill = as.factor(group), alpha = 0.5),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=group, 
                color = as.factor(group)), size = 1)+
  facet_wrap(~group, 
             scales = "free_y",  
             nrow = 4)   +
  scale_fill_viridis_d(direction = 1) +
  scale_color_viridis_d(direction = 1)

```

```{r data-heatmap-moy-group}

data_group <- data_pick %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id"))

data_heatmap_moy_group <- quantile_gran(data_group,
                                  gran1="month_year",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")

moy_group <- data_heatmap_moy_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 4) +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
  ylab("demand (in Kwh)")   +
  scale_fill_viridis_d() +
  scale_color_viridis_d()
```

```{r data-heatmap-wkndwday-2}

data_heatmap_wkndwday <- quantile_gran(data_pick, "wknd_wday", quantile_prob_val = c(0.25, 0.5, 0.75)) %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_heatmap_wkndwday$category <- factor(data_heatmap_wkndwday$category, levels = c("Weekday", "Weekend"))

p1 <- data_heatmap_wkndwday %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
  left_join(cluster_result, by = c("customer_id" = "id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=customer_id), fill = "lightblue") +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 6)  +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

```

```{r data-heatmap-wkndwday-group}
data_group <- data_pick %>%
  mutate(customer_id = as.character(customer_id)) %>%
  left_join(cluster_result, by = c("customer_id" = "id"))

wkndwday_data<- data_group %>% create_gran("wknd_wday") 
ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]

wkndwday_group <- wkndwday_data%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  lvplot::geom_lv(aes(fill = as.factor(group), color = as.factor(group)), k=5, alpha = 0.5) +
  #geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             nrow = 4)   +
  scale_fill_viridis_d() +
  scale_color_viridis_d()

```

```{r combined, eval = FALSE}

combined <- hod_group + moy_group + wkndwday_group & theme(legend.position = "bottom")*theme_characterisation()
combined + plot_layout(guides = "collect")

```