### 2021-10-27 selecting_customers.Rmd
adding prototypes

---
title: "Plots with 5 anchors"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsr)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
library(patchwork)
library(GGally)
library(distributional)
library(viridis)
library(Rtsne)
library(liminal)
```

```{r mytheme}
theme_characterisation <- function() {

  theme_light() + # setting theme
    #theme(strip.text = element_text(margin = margin(b = 0, t = 0))) + # narrow facet space
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + # no axis ticks 
  theme(panel.spacing =unit(0, "lines")) +  # to ensure no gap between facets
    theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) + # no x-axis labels to further reduce the gap between facets
    #theme(axis.text.x = element_text(angle=90, hjust=1, size = 9)) + # rotate the x-axis text
  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
  #theme(axis.text.x = element_text(size=5)) +
    theme(strip.background = element_blank(),
  strip.text.x = element_blank())
}
```

_Set number of anchors_
```{r anchor}
kopt <- 6# number of clusters
```

_Load the customers you want to cluster_
```{r}
quantile_prob_graph <- c(0.25, 0.5, 0.75)

# data_pick_one <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
# data_pick_two <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
# data_pick_three <- c(8221762, 8273636, 10359424, 8232822)%>% as_tibble %>% set_names("customer_id")

data_pick_one <- c(8541744, 9355808, 8603880, 8619309, 10542667) %>% as_tibble %>% set_names("customer_id")
#data_pick_two <- c(8688242, 8643837, 8184707, 10534355, 8684420) %>% as_tibble%>% set_names("customer_id")
data_pick_three <- c(9792072, 8589936, 8454235, 10692366, 8603828)%>% as_tibble%>% set_names("customer_id")
data_pick_four <- c(8618759, 8291696, 10357256, 8290374) %>% as_tibble %>% set_names("customer_id")
data_pick_five <- c(9044864, 8642053, 10534367, 9021526,11162275) %>% as_tibble %>% set_names("customer_id")
data_pick_six <- c(8221762, 8273636, 10359424, 8232822, 11450499)%>% as_tibble %>% set_names("customer_id")

```

```{r assemble}
data_pick_cust <- bind_rows(
data_pick_one, 
#data_pick_two, 
data_pick_three,
data_pick_four,
data_pick_five, 
data_pick_six,
.id = "design") %>% 
  mutate(customer_id = as.character(customer_id))

```

```{r tsne, eval = FALSE}
data_356cust_wide <- read_rds("data/data_356cust_wide.rds")

data_356cust_pc <- prcomp(data_356cust_wide[,-1],
                          center = FALSE, scale = FALSE, retx = TRUE)
set.seed(2935)
data_356cust_pc10 <- as_tibble(data_356cust_pc$x[,1:6])

##---- t-SNE embeddings

library(Rtsne)
set.seed(2099)

tSNE_fit <- data_356cust_wide%>% 
  select(-customer_id) %>% 
         Rtsne(PCA = FALSE,
         perplexity = 30)

tsne_df <- data.frame(tsneX = tSNE_fit$Y[, 1], tsneY = tSNE_fit$Y[, 2], customer_id = as.character(data_356cust_wide$customer_id)) %>% 
  left_join(
  data_pick_cust, by = c("customer_id"
)) %>% mutate(design = if_else(is.na(design), "0", design))

#rownames(tsne_df) <- data_356cust_wide$customer_id

rownames(data_356cust_pc10) <- data_356cust_wide$customer_id

## ----tsne-xy--------------------------------------------------------
tsne_xy <- ggplot(tsne_df, aes(x = tsneX, y = tsneY, color = design)) +
  geom_point(aes(text = customer_id)) +
  #scale_color_manual(values = limn_pal_tableau10()) +
  scale_colour_viridis_d(direction = -1) +
  guides(color = FALSE) +
  labs(caption = "tSNE") +
  theme(aspect.ratio = 1) +
  theme_light()

## ----highlight-customer--------------------------------------------------------

library(plotly)
tsne_plotly <- tsne_xy %>% ggplotly(tooltip = "text")
tsne_plotly

```

_Load raw data for these customers and scale it according to nqt method and group them_

```{r}
data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
  mutate(customer_id = as.character(customer_id)) %>% 
  filter(customer_id %in% data_pick_cust$customer_id) %>% 
  gracsr::scale_gran( method = "nqt",
                      response = "general_supply_kwh")
```

```{r clustering}

hod <- data_pick %>% 
  dist_gran(gran1 = "hour_day", response = "general_supply_kwh")

moy <- data_pick %>% 
  dist_gran(gran1 = "month_year", response = "general_supply_kwh")

wkndwday <- data_pick %>% 
  dist_gran(gran1 = "wknd_wday", response = "general_supply_kwh")

distance <- wkndwday/2 + moy/12 + hod/24

f = as.dist(distance)

cluster_result <- f %>% clust_gran(kopt = 5) %>% 
  rename("customer_id" = "id")
```

```{r hod-data}
data_hod <- quantile_gran(data_pick,
                                  "hour_day",
                                  quantile_prob_val = quantile_prob_graph) %>% 
  pivot_wider(names_from = quantiles,
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id")) %>% 
  left_join(cluster_result, by = c("customer_id"))

#data_heatmap_hod$customer_id = factor(data_heatmap_hod$customer_id, levels = data_pick_cust$value)
data_hod$category <- factor(data_hod$category, levels = 0:23)

# lev <- data_hod %>% distinct(customer_id, design) %>% arrange(design) %>% pull(customer_id)
```

```{r hod-ind-design}

# data_hod$customer_id <- factor(data_hod$customer_id, levels = lev)

hod_ind_design <- data_hod %>% 
  left_join(cluster_result,by = c("customer_id", "group")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, 
                  fill = design),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id,
                color = design), 
            size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y",
             ncol=24) + theme_characterisation()+
  xlab("hour-of-day") +
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  scale_x_discrete(breaks = seq(0, 23, 3))
```

```{r hod-ind-group}
hod_ind_group <- data_hod %>% 
  left_join(cluster_result) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=customer_id, fill = as.factor(group)),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y",
              ncol=24) + theme_characterisation()+
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) +    xlab("hour-of-day")  +
  scale_x_discrete(breaks = seq(0, 23, 3))
# + scale_y_continuous(breaks = NULL) 
# 
# 
(hod_ind_design/hod_ind_group)* theme(legend.position = "none")

```

```{r data-heatmap-moy-design}

data_moy <- quantile_gran(data_pick,
                                  "month_year", 
                                  quantile_prob_val = quantile_prob_graph) %>% 
  pivot_wider(names_from = quantiles, 
              values_from = quantiles_values) %>% 
  left_join(data_pick_cust, by = c("customer_id"))

data_moy$category <- factor(data_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

moy_ind_design <- data_moy %>% 
  left_join(cluster_result, by = c("customer_id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(design)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(design)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             ncol = 24) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_viridis_d(direction = 1)+
  scale_color_viridis_d(direction = 1)
```

```{r data-heatmap-moy-group-ind}

moy_ind_group <- data_moy %>% 
  left_join(cluster_result, by = c("customer_id")) %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, 
                  group=customer_id, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=customer_id, color = as.factor(group)), size = 1) +
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             ncol = 24) +
    ylab("demand (in Kwh)") +
    xlab("month-of-year")  +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))

(moy_ind_design/moy_ind_group)*theme(legend.position = "none")

```

```{r data-heatmap-wkndwday-group}

data_wkndwday <- data_pick  %>%
  left_join(cluster_result)%>% create_gran("wknd_wday")  %>% 
  left_join(data_pick_cust, by = c("customer_id"))

ylim1 = boxplot.stats(data_wkndwday$general_supply_kwh)$stats[c(1, 5)]

wkndwday_ind_design <- data_wkndwday%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  #lvplot::geom_lv(aes(fill = as.factor(design), 
   #                   color = as.factor(design)), k=5, alpha = 0.5) +
  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
              ncol = 24)  +
  theme_characterisation() +
  scale_fill_viridis_d(direction = 1)+
  scale_color_viridis_d(direction = 1)

```

```{r}
wkndwday_ind_group <- data_wkndwday%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  #lvplot::geom_lv(aes(fill = as.factor(group), 
          #            color = as.factor(group)), k=5, alpha = 0.5) +
  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05)+
  facet_wrap(design~customer_id, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 24)  +
  theme_characterisation() +
  scale_fill_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00",  "grey", "black"))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))

(wkndwday_ind_design / wkndwday_ind_group)* theme(legend.position = "none")
```

# contribution of each levels

```{r}

```

Draw group ones and make interpretation:

```{r}
data_validation <- (hod/24) %>% broom::tidy() %>% 
  rename("hod" = "distance") %>% 
  left_join((moy/12) %>% 
              broom::tidy(), 
            by = c("item1", "item2")) %>% 
  rename("moy" = "distance") %>% 
  left_join((wkndwday/2) %>% 
              broom::tidy(), 
            by = c("item1", "item2"))%>% 
  rename("wkndwday" = "distance") %>% 
  left_join(cluster_result, by = c("item1" = "customer_id")) %>% 
  rename("group_item1" = "group") %>% 
  left_join(cluster_result, by = c("item2" = "customer_id")) %>%  
  rename("group_item2" = "group") %>% 
  pivot_longer(3:5,names_to="gran",
               values_to = "distance")

data_validation %>%
  group_by(gran) %>%
  summarise(d = sd(distance))

data_validation %>%
  group_by(gran, group_item1, group_item2) %>% 
  summarise(sum = sum(distance)) %>% 
  pivot_wider(names_from = group_item2, values_from = sum) %>% 
  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`)) %>% 
  select(-(3:7)) %>% 
  pivot_wider(names_from = group_item1, values_from = distance)

```

```{r}

hod_cat <- data_pick %>% 
  dist_gran_cat(gran1 = "hour_day", response = "general_supply_kwh") %>% 
  mutate(gran = "hod")

moy_cat <- data_pick %>% 
  dist_gran_cat(gran1 = "month_year", response = "general_supply_kwh")%>%
  mutate(gran = "moy")

wkndwday_cat <- data_pick %>% 
  dist_gran_cat(gran1 = "wknd_wday", response = "general_supply_kwh")%>% 
  mutate(gran = "wnwd")

data_validation_cat <- bind_rows(hod_cat, moy_cat, wkndwday_cat)

gran_cat_dist <- data_validation_cat %>% 
  left_join(cluster_result, by = c("customer_from" = "customer_id")) %>% 
  rename("group_from" = "group") %>% 
  left_join(cluster_result, by = c("customer_to" = "customer_id")) %>%  
  rename("group_to" = "group") %>% 
  group_by(gran,
           category,
           group_from, 
           group_to) %>% 
  summarise(sum = sum(distance)) %>% 
  pivot_wider(names_from = group_to, values_from = sum) %>% 
  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`)) %>% 
  select(-(4:8)) %>% 
  arrange(-distance)

```

# Contribution of individual category for hod

```{r}
gran_cat_hod <-   gran_cat_dist %>% 
  filter(gran=="hod")

gran_cat_hod$category <- factor(gran_cat_hod$category, levels = 0:23)

# for each group, percent contribution of hours

group_total <- gran_cat_hod %>% 
  ungroup() %>% 
    group_by(group_from) %>% 
    summarise(group_total = sum(distance))

  # gran_cat_hod %>% 
  #   ungroup %>% 
  #   left_join(group_total, by = "group_from") %>% 
  #   mutate(cat_contibution = distance*100/group_total) %>% 
  #   group_by(group_from) %>% 
  #   slice(c(1:3))

    gran_cat_hod %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% slice(c(1:4)) %>% 
      ggplot(aes(group_from, category, color = as.factor(group_from))) +
      geom_point(aes(size = distance))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()

data_pcp <- gran_cat_hod %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% 
  select(group_from, category, cat_contibution) %>% 
  pivot_wider(names_from = "category", values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))

parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 2:ncol(data_pcp),
                   groupColumn = "group_from",
                   showPoints = TRUE, 
                   alphaLines = 1,
                   order = "anyClass",
                   scale = "globalminmax"
) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) + theme_bw()
    parcoord

```

# Contribution of individual category for moy

```{r}
gran_cat_moy <-   gran_cat_dist %>% 
  filter(gran=="moy")

gran_cat_moy$category <- factor(gran_cat_moy$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

group_total <- gran_cat_moy %>% 
  ungroup() %>% 
    group_by(group_from) %>% 
    summarise(group_total = sum(distance))

    gran_cat_moy %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% slice(c(1:3)) %>% 
      ggplot(aes(group_from, category, color = as.factor(group_from))) +
      geom_point(aes(size = distance))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()

    data_pcp <- gran_cat_moy %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% 
  select(group_from, category, cat_contibution) %>% 
  pivot_wider(names_from = "category", values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))

parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 2:ncol(data_pcp),
                   groupColumn = "group_from",
                   showPoints = TRUE, 
                   alphaLines = 1,
                   order = "anyClass",
                   scale = "globalminmax"
) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black")) + theme_bw()
    parcoord

```

# Contribution of individual category for wknwday

```{r}
gran_cat_wkndwday <-   gran_cat_dist %>% 
  filter(gran=="wnwd")

gran_cat_wkndwday$category <- factor(gran_cat_wkndwday$category, levels = c("Weekday", "Weekend"))

group_total <- gran_cat_wkndwday %>% 
  ungroup() %>% 
    group_by(group_from) %>% 
    summarise(group_total = sum(distance))

    gran_cat_wkndwday %>% 
    ungroup %>% 
    left_join(group_total, by = "group_from") %>% 
    mutate(cat_contibution = distance*100/group_total) %>% 
    group_by(group_from) %>% 
    arrange(-distance) %>% slice(c(1)) %>% 
      ggplot(aes(group_from, category, color = as.factor(group_from))) +
      geom_point(aes(size = distance))+
  scale_color_manual(values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey", "black"))+ theme_light()
```

# Show aggregate group behavior

```{r data-heatmap-hod-group}
legend_title <- "group"

data_group <- data_pick  %>% 
  left_join(cluster_result, by = c("customer_id"))

data_heatmap_hod_group <- quantile_gran(data_group,
                                  gran1="hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)

# data_heatmap_hod_group <- data_heatmap_hod_group %>%
#   mutate(group = case_when(group==1 ~"Group 2",
#                            group==2 ~"Group 4",
#                            group==3 ~"Group 1",
#                            group==4 ~"Group 3"))

# data_heatmap_hod_group$group <- factor(data_heatmap_hod_group$group, levels = c("Group 1", "Group 2", "Group 3", "Group 4"))

data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")

hod_group <- data_heatmap_hod_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=group,
                  fill = as.factor(group), alpha = 0.5),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=group, 
                color = as.factor(group)), size = 1)+
  facet_wrap(~group, 
             scales = "free_y",  
             ncol = 5) + 
              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
  ylab("demand (in Kwh)") + 
  theme_bw() +
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  theme(panel.spacing =unit(0, "lines")) + 
  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7)) +
  scale_x_discrete(breaks = seq(1, 24, 3))+ 
  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
  theme(plot.margin = margin(0, 0, 0, 0, "cm")) +
   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
  theme(legend.position = "bottom")

```

```{r data-heatmap-moy-group}
data_heatmap_moy_group <- quantile_gran(data_group,
                                  gran1="month_year",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")

moy_group <- data_heatmap_moy_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 5) +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
  ylab("demand (in Kwh)") +
  theme_bw() +
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + 
  theme(panel.spacing =unit(0, "lines")) + 
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
   scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) + 
  theme(legend.position = "bottom") +
  theme(panel.spacing =unit(0, "lines")) + 
  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
```

```{r}
combined <- (hod_group/moy_group) & theme(legend.position = "bottom")
combined + plot_layout(guides = "collect")
```

```{r}
wkndwday_data <- data_group %>% create_gran("wknd_wday") %>% 
  create_gran("hour_day")

ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]

wkndwday_group <- wkndwday_data%>% 
  ggplot(aes(x=hour_day, y = general_supply_kwh)) +
  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
  geom_boxplot(aes(fill = as.factor(group), color = as.factor(group)),alpha = 0.5)+
  #geom_boxplot(outlier.size = 1) + 
  coord_cartesian(ylim = ylim1*1.05)+
  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
#geom_boxplot(aes(fill = as.factor(group))) +
  #scale_fill_lv() +
 xlab("wknd-wday") + 
  ylab("demand (in Kwh)") +
   facet_grid(wknd_wday~group, 
             scales = "free_y", 
             labeller = "label_both") + 
  theme_bw() +
   theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + 
  theme(panel.spacing =unit(0, "lines")) + 
  theme(plot.margin = margin(0, 0, 0, 0, "cm"))  +
  scale_fill_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  scale_color_manual(legend_title, values = c("#E69F00", "#009E73","#0072B2", "#D55E00", "grey")) +
  theme(legend.position = "bottom") +
  theme(panel.spacing =unit(0, "lines")) + 
  theme(axis.text.x = element_text(angle=90, hjust=1, size = 7))
wkndwday_group
```