### 2021-11-02 paper-supplementary.Rmd
proof reading changes

---
title: "<small> Supplementary materials for the main submission entitled - <br><br> Clustering time series based on probability distributions across temporal granularities"

bibliography: [bibliography.bib]
preamble: >
  \usepackage{mathtools,amssymb,booktabs,amsthm,todonotes,colortbl}
  \def\mod{~\text{mod}~}
  \newtheorem{definition}{Definition}
  \usepackage{mathptmx}
  \usepackage{caption}
  \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
  \captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
  \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
  \def\novspacing{\setlength{\aboverulesep}{0pt}\setlength{\belowrulesep}{0pt}}
  \def\vspacing{\setlength{\aboverulesep}{0.4ex}\setlength{\belowrulesep}{0.65ex}}
output:
  bookdown::pdf_book:
    MonashEBSTemplates::workingpaper:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
---

```{r external, include = FALSE, cache = FALSE}
knitr::read_chunk('script/main.R')
```

```{r load-lib}

```

```{r mytheme-application}
```

```{r theme_characterisation}
```

```{r theme-validation}

```

## Raw time series plot

\noindent To get a sense of how the raw time series data looks, we plot the energy usage for $50$ sampled households is plotted along the y-axis versus time from past to future in Figure ref\{fig:raw-data-50}. Each of these series is associated with a single customer. Energy consumption for each customer is given at fine temporal resolution (every 30 minutes) for a period of 2-3 years.

```{r raw-data-50, out.width = "100%", fig.cap="The raw data for 50 households are shown. It looks like there is a lot of missing values and unequal length of time series along with asynchronous periods for which data is observed. No insightful behavioral pattern could be discerned from this view other than when the customer is not at home."}

knitr::include_graphics("figs/raw_plot_cust.png") # look at smart-meter.R for the code
```

## Plot displaying missing observations

_Prototype selection method_

S1. Robust scaling is applied to each customer.

S2. $50^{th}$ percentile for each category for each granularity is obtained for each customers. So we have a data structure with  $356$ rows and $(24 + 12 + 2)$ variables corresponding to $50^{th}$ percentile for each hour-of-day, month-of-year and weekend-weekday.

S3. Apply principal components and restrict the results down to the first six principal components (which makes up approximately 85% of the variance explained in the data) to use with the grand tour.

S4. Run t-SNE using the default arguments on the complete data (sets the perplexity to equal 30 and performs random initialisation). We then create a linked tour with t-SNE layout with liminal as shown in Figure 4.

S5. We inspect of the subspace generated by the set of low-dimensional projections in tour by looking for a simplex shape while the visualization moves from one basis to another. When we brush the corners of the simplex, we find they fall on the edge of the t-SNE point cloud.
Hall, Marron, and Neeman (2005) have shown that in the extreme case of high-dimension, low-sample size data, observations are on the vertices of a simplex.

This is because in high-dimensional data analysis the curse of dimensionality reasons that points tend to be far away from the center of the distribution and on the edge of high-dimensional space. Contrary to this, is that projected data tends to clump at the center.

S6. These points should ideally correspond to different behavior with respect to all the variables considered while running PCA.

```{r mds, out.width="100%", fig.cap="Instance selection using tours and projecting the points in a lower dimensional tsne cloud."}
knitr::include_graphics("figs/liminal_view.png")
```

<!-- From this set we select $4$ "anchor" customers which are far apart from each other and $5$ neighboring customers for each of these anchors. These selections were done using the granularity $hod$ space. It is important to note that when we use our proposed methodologies, it is based on all dimensions `hod`, `moy` and `wkndwday`.Fig \ref{fig:mds} shows the MDS of these 356 customers in a 2D space basis their distance on individual granularities and when all of them are combined. Our methodolgies are run on these $24$ customers, which act as a way to evaluate the proposed methodologies.  -->

<!-- . Instances may be distributed in the representation space in a reasonable manner, revealing their similarities, using this instance wise discriminative learning. A dual-level progressive similar instance selection (DPSIS) approach could also be used -->

## Clustering on 350 customers

```{r data-load-entire}
# Read the nqt distances

wkndwday <- read_rds(here("data/dist_gran_wkndwday_356cust_nqt.rds")) %>% broom::tidy()

moy <- read_rds(here("data/dist_gran_moy_356cust_nqt.rds")) %>% broom::tidy()

hod <- read_rds(here("data/dist_gran_hod_356cust_nqt.rds")) %>% broom::tidy()

# Make the distance metrics

distance <- wkndwday %>% 
  left_join(moy, by = c("item1", "item2")) %>% 
  left_join(hod, by = c("item1", "item2")) %>% 
  rename("wkndwday" ="distance.x",
         "moy" = "distance.y",
         "hod" = "distance") %>%
  mutate(item1 = as.integer(as.character(item1)),
         item2 = as.integer(as.character(item2)))

filtered_distance <- distance %>%
  filter(!(item1 %in% c(8196183, 8508008, 8680538))) %>% 
  filter(!(item2 %in% c(8196183, 8508008, 8680538)))

total_distance <- filtered_distance %>% 
  mutate(total = wkndwday/2 + moy/12 + hod/24)

total_distance_wide <- total_distance %>% pivot_wider(-c(2:5), 
                                                      names_from = item2,
                                                      values_from = total)

rownames(total_distance_wide) <- total_distance_wide$item1

mds_data <- total_distance_wide %>% 
  mutate_all(~replace(., is.na(.), 0)) %>%
  tibble::rownames_to_column() %>%  
  dplyr::select(-item1) %>% 
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from=rowname, values_from=value)

rownames(mds_data) <- total_distance_wide$item1

df <- mds_data[-1] %>% as.matrix()
DM <- matrix(0, ncol(mds_data), ncol(mds_data))
DM[lower.tri(DM)] = df[lower.tri(df, diag=TRUE)] # distance metric
f = as.dist(DM)

first_lot <- mds_data %>% names()

id <- c(first_lot[-1], mds_data$name[nrow(mds_data)])

# Find optimal number of clusters

library(fpc)
library(cluster)
k = array()
for(i in 5:50)
{
  group <- f %>% hclust (method = "ward.D") %>% cutree(k=i)
  p <- cluster.stats(f, clustering = group, silhouette = TRUE)
  k[i]=p$sindex
}

ggplot(k %>% as_tibble %>% mutate(k = row_number()), aes(x=k, y = value)) + geom_line() + scale_x_continuous(breaks = seq(2, 50, 2))

# plot(k, type = "l")
# 6 coming as the number of clusters with maximum silwidth

group <- f %>% hclust (method = "ward.D") %>% cutree(k=17)
cluster_result <- bind_cols(customer_id = id, group = group)

# cluster_result %>% group_by(group) %>% count()

```

```{r data-heatmap-hod-group-entire}
legend_title <- "group"

data_pick <- read_rds(here::here("data/elec_nogap_2013_clean_356cust.rds")) %>%
  mutate(customer_id = as.character(customer_id)) %>%
  gracsr::scale_gran( method = "robust",
                      response = "general_supply_kwh")

data_group <- data_pick  %>% 
  mutate(customer_id = as.character(customer_id)) %>% 
   gracsr::scale_gran( method = "robust",
                       response = "general_supply_kwh") %>% 
  left_join(cluster_result, by = c("customer_id"))

data_heatmap_hod_group <- quantile_gran(data_group,
                                  gran1="hour_day",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_hod_group$category <- factor(data_heatmap_hod_group$category, levels = 0:23)

data_heatmap_hod_group$group <- paste("group", data_heatmap_hod_group$group, sep = "-")

hod_group_entire <- data_heatmap_hod_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`,
                  group=group,
                  fill = as.factor(group), alpha = 0.5),
              alpha = 0.5) +
  geom_line(aes(y = `50%`,
                group=group, 
                color = as.factor(group)), size = 1)+
  facet_wrap(~group, 
             scales = "free_y",  
             ncol = 17) + 
              #labeller = labeller(xfacet = c(`1` = "Group 2", `2` = "Group 4",`3` = "Group 1",`4` = "Group 3"))
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("hour-of-day") + 
  ylab("demand (in Kwh)") + 
  theme_bw()  +
  scale_x_discrete(breaks = seq(1, 24, 3))+ 
  #theme(strip.text = element_text(size = 8, margin = margin(b = 0, t = 0)))+
  theme_application() +
  scale_fill_manual(values=as.vector(polychrome(20)))+
  scale_color_manual(values=as.vector(polychrome(20)))+
  theme(legend.position = "bottom")

```

```{r data-heatmap-moy-group-entire}
data_heatmap_moy_group <- quantile_gran(data_group,
                                  gran1="month_year",
                                  quantile_prob_val = c(0.25, 0.5, 0.75),
                                  group="group") %>% 
  pivot_wider(names_from = quantiles, values_from = quantiles_values)

data_heatmap_moy_group$category <- factor(data_heatmap_moy_group$category, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

data_heatmap_moy_group$group <- paste("group", data_heatmap_moy_group$group, sep = "-")

moy_group_entire <- data_heatmap_moy_group %>% 
  ggplot(aes(x = category)) + 
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`, group=group, fill = as.factor(group)), alpha = 0.5) +
  geom_line(aes(y = `50%`, group=group, color = as.factor(group)), size = 1 ) +
  facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_value",
             ncol = 17) +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0))) + xlab("month-of-year") + 
  ylab("demand (in Kwh)") +
  theme_bw() + theme_application() +
  scale_fill_manual(values=as.vector(polychrome(20)))+
  scale_color_manual(values=as.vector(polychrome(20)))+
  theme(legend.position = "bottom")
```

```{r data-wnwd-group-entire}
wkndwday_data <- data_group %>%
  create_gran("wknd_wday") 
wkndwday_data$group <- as.factor(wkndwday_data$group)
# %>% 
#   create_gran("hour_day")

ylim1 = boxplot.stats(wkndwday_data$general_supply_kwh)$stats[c(1, 5)]

wkndwday_group_entire <- wkndwday_data%>% 
  ggplot(aes(x=wknd_wday, y = general_supply_kwh)) +
  #lvplot::geom_lv(aes(fill = as.factor(group)), k=5) +
  geom_boxplot(aes(fill = group, color = group),alpha = 0.5, outlier.alpha = 0.05)+
  #geom_boxplot(outlier.size = 1) + 
  coord_cartesian(ylim = ylim1*1.05)+
  #ggridges::geom_density_ridges2(aes(x = general_supply_kwh, y = wknd_wday,fill = as.factor(group))) + coord_flip() +
#geom_boxplot(aes(fill = as.factor(group))) +
  #scale_fill_lv() +
 xlab("wknd-wday") + 
  ylab("demand (in Kwh)") +
   facet_wrap(~group, 
             scales = "free_y", 
             labeller = "label_both",
             ncol = 17) + 
  theme_bw() + theme_application() +
  scale_fill_manual(values=as.vector(polychrome(20)))+
  scale_color_manual(values=as.vector(polychrome(20)))+
  theme(legend.position = "none")
```

```{r combined-groups-js-entire, fig.cap = "The distribution of electricity demand for the clusters across hod (a), moy (b) and wkndwday (c). It seems like group 2 and 5 have a hod pattern across its members, while group 1, 3, 5 have a moy pattern. Wknd-wday variations across groups are not distinguishable, indicating that it is not a critical variable for clustering. It is helpful to compare the summarised distributions of groups to that of individuals to confirm that the most of individuals in the group have the same characterisation."}

 (hod_group_entire/moy_group_entire/ wkndwday_group_entire) +
  plot_annotation(tag_levels = 'a', tag_prefix = '(', tag_suffix = ')')+
  plot_layout(guides = "collect")& theme(legend.position = 'none')
```

# contribution

```{r contribution}
# data_validation <- hod %>%  
#   rename("hod" = "distance") %>% 
#   left_join(moy, 
#             by = c("item1", "item2"))%>% 
#   rename("moy" = "distance") %>% 
#   left_join((wkndwday), 
#             by = c("item1", "item2"))%>% 
#   rename("wkndwday" = "distance") %>% 
#   left_join(cluster_result, by = c("item1" = "customer_id")) %>% 
#   rename("group_item1" = "group") %>% 
#   left_join(cluster_result, by = c("item2" = "customer_id")) %>%  
#   rename("group_item2" = "group") %>% 
#   mutate(hod = hod/24, moy = moy/12, wkndwday = wkndwday/2) %>% 
#   filter(!is.na(group_item1)| !is.na(group_item2)) %>% 
#   pivot_longer(3:5,names_to="gran",
#                values_to = "distance")  
#

data_validation_cat <- read_rds(here::here("validation/data_validation_cat_all_cluster.rds"))

# hod_cat <- data_pick %>%
#   dist_gran_cat(gran1 = "hour_day", response = "general_supply_kwh") %>%
#   mutate(gran = "hod")
# 
# moy_cat <- data_pick %>%
#   dist_gran_cat(gran1 = "month_year", response = "general_supply_kwh")%>%
#   mutate(gran = "moy")
# 
# wkndwday_cat <- data_pick %>%
#   dist_gran_cat(gran1 = "wknd_wday", response = "general_supply_kwh")%>%
#   mutate(gran = "wnwd")
# 
# 
# data_validation_cat <- bind_rows(hod_cat, moy_cat, wkndwday_cat) %>%
#   mutate(customer_from = as.character(customer_from),
#          customer_to = as.character(customer_to))

gran_cat_dist <- data_validation_cat %>% 
  mutate(customer_from = as.character(customer_from),
         customer_to = as.character(customer_to)) %>% 
  left_join(cluster_result, by = c("customer_from" = "customer_id")) %>% 
  rename("group_from" = "group") %>% 
  left_join(cluster_result, by = c("customer_to" = "customer_id")) %>%  
  rename("group_to" = "group") %>% 
  group_by(gran,
           category,
           group_from, 
           group_to) %>% 
  summarise(sum = sum(distance)) %>% 
  pivot_wider(names_from = group_to, values_from = sum) %>% 
  mutate(distance = sum(`1`, `2`, `3`, `4`, `5`,`6`, `7`, `8`, `9`, `10`, `11`, `12`, `13`, `14`, `15`, `16`, `17`)) %>% 
  select(-(4:8)) %>% 
  arrange(-distance)

# Contribution of individual category for hod

gran_cat_hod <-   gran_cat_dist %>% 
  filter(gran=="hod") %>% 
  mutate(category = as.integer(category)) %>% 
  arrange(group_from, category)

gran_cat_hod$category <- factor(gran_cat_hod$category, levels = 0:23)

# for each group, percent contribution of hours

group_total <- gran_cat_hod %>% 
  ungroup() %>% 
  group_by(group_from) %>% 
  summarise(group_total = sum(distance))

data_pcp <- gran_cat_hod %>% 
  ungroup %>% 
  left_join(group_total, by = "group_from") %>% 
  mutate(cat_contibution = distance*100/group_total) %>% 
  group_by(group_from) %>% 
  #arrange(-distance) %>% 
  select(group_from, category, cat_contibution) %>% 
  pivot_wider(names_from = "category", 
              values_from = "cat_contibution") %>% mutate(group_from = as.factor(group_from))

parcoord <- GGally::ggparcoord(data_pcp ,
                               columns = 2:ncol(data_pcp),
                               groupColumn = "group_from",
                               showPoints = TRUE, 
                               alphaLines = 1,
                               scale = "globalminmax"
) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")+
  scale_fill_manual(values=as.vector(polychrome(20)))+
  theme_bw()
parcoord 
```