### 2021-11-07 supplementary/old/analysis.Rmd

---
title: "analysis"
author: "Sayani Gupta"
date: "05/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-lib}
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
```

```{r load-data}
load("data/ALL_DATA.Rdata")
```

```{r miss-data}
empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    ifelse(as.character(x)!="", x, NA)
}

## transform all columns
CustomerData <- CustomerData %>% mutate_each(funs(empty_as_na))

vis_miss(CustomerData, warn_large_data = FALSE) + 
  theme(axis.text.x = element_text(angle = 90, size = 5))
```

```{r miss-var}
gg_miss_var(CustomerData) + theme_bw() +
  theme(axis.text.y = element_text(size = 4))
```