### 2021-11-07 supplementary/old/introduction.Rmd

---
title: "introduction_section"
author: "Sayani Gupta"
date: "08/09/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Paper: Clustering of residential electricity customers using load time series

- Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal time-series common in residential datasets.

 - In this paper we propose a strategy to alleviate these limitations by
converting any types of load time series into map models that can be readily clustered.

- higher cluster distinction and robustness

- The experiment with 12 clusters results in around 61% distinction, improved coincidence
factor by around 6.75% relative to a random grouping, and robustness of around 59% against the applied noise.

- stochastic and not deterministic segmentation

- The processes of stacking daily
curves and applying PCA to extract core features from the pile of
curves, do not suit parallel computing. This makes the algorithm unsuitable for a large number of customers.

-  Alternatively, Shannonian entropy [62,63], can be used to optimise the embedding
parameters.

- feature based sensitive to outlier days

- The method feature based  is static in that any
addition of new data (of additional customers and or additional days)
require repeating both feature extraction and clustering

- model based : Regardless of its length, dates and times, each
continuous series translates to a map function or yields a geometry
 Alternatively,
Shannonian entropy [62,63], can be used to optimise the embedding
parameters.

 A distinction metric is the silhouette width [65,50], which measures how
similar homes are to their own cluster members (cohesion) compared
with homes in other clusters (separation), and is bounded within (−1,
1).

This is a desirable outcome; and clustering methods which are
able to isolate outliers are found to exhibit high performance as indicated in

Though, with a finite set of clusters the outliers stand
out, forcing the rest of the samples to lump together in big clusters with
high variability and vague profiles. To remedy this situation, we can
exclude outliers and re-cluster samples.

To remedy this situation, we can
exclude outliers and re-cluster samples. Though exclusion of outliers
often gives way to new outliers to emerge. Alternatively, we reduce
clusters to between 8 and 12 to retain all samples. The optima are
obtained at 9 and 12 clusters with ∼62.4% and ∼61.1% distinctions,
respectively.

There are at least two key areas to extend this research. These include load prediction based on clustering repeatability, and customer
privacy protection by noise encryption. The current repeatability metric
determines if the same clusters of homes are formed by clustering over
successive periods of time, such as successive seasons, or years. To reinforce repeatability, future work can also focus on cross validation by
bootstrapping, i.e., the same clusters of homes should be formed based
on randomly selected days within a certain period.